
---
title: "Main R Analysis Runner"
output: rmarkdown::notebook
---

## Required R Packages

This document requires R 4.0+ and the following packages:
- **readr** — read CSV files
- **dplyr** — data manipulation
- **tidyr** — data reshaping
- **ggplot2** — visualization
- **lme4, lmerTest** — mixed-effects models
- **emmeans** — estimated marginal means
- **car** — model diagnostics
- **rempsyc, flextable** — table formatting
- **rmarkdown** — document rendering

**First-time setup (run once in R/RStudio):**
```r
packages <- c("readr", "dplyr", "tidyr", "ggplot2", "lme4", "lmerTest", 
              "emmeans", "car", "rempsyc", "flextable", "rmarkdown")
install.packages(packages)
```

## Automatic Package Management and Document Rendering

This chunk automatically loads required R packages and then renders the analysis notebooks.

```{r setup, include=FALSE}
# Automatic package management for pons.lea
packages <- c(
  "readr", "dplyr", "tidyr", "ggplot2",
  "lme4", "lmerTest", "emmeans", "car",
  "rempsyc", "flextable", "knitr"
)

# Install missing packages (commented by default for safety)
# install.packages(setdiff(packages, rownames(installed.packages())))

# Load all packages
invisible(lapply(packages, library, character.only = TRUE))

cat("✓ All required R packages loaded\n")
```

```{r render_notebooks, include=TRUE}
# Render all analysis notebooks in sequence
notebooks_to_render <- c(
  "notebooks/q_preprocessing.Rmd",
  "notebooks/movement_statistics.Rmd",
  "notebooks/q_statistics.Rmd",
  "notebooks/movement_first_impression.Rmd"
)

rendered_outputs <- character(length(notebooks_to_render))

for (i in seq_along(notebooks_to_render)) {
  rendered_outputs[i] <- rmarkdown::render(
    notebooks_to_render[i],
    envir = new.env(parent = globalenv()),
    quiet = TRUE
  )
}

data.frame(
  notebook = notebooks_to_render,
  output = rendered_outputs
)
```
