---
title: "Run full R pipeline"
subtitle: "Questionnaire merge + statistical analysis"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

# Renders a .Rmd, then deletes the "<name>_files/" and "<name>_cache/"
# folders knitr creates next to it (embedded HTML figures, and cached chunk
# results - statistical_analysis.Rmd's "robustness" chunk has cache=TRUE).
# The real results are already saved properly by save_figure()/save_table()
# into results/, so both folders are just clutter in code/.
#
# Note: deleting the cache means the "robustness" chunk (the cluster
# bootstrap, B = 1000 x 3 models) is recomputed in full on every run instead
# of being skipped when unchanged - i.e. every run is as slow as the first.
render_and_clean <- function(rmd_path, ...) {
  rmarkdown::render(rmd_path, ...)
  base <- file.path(dirname(rmd_path), tools::file_path_sans_ext(basename(rmd_path)))
  for (suffix in c("_files", "_cache")) {
    d <- paste0(base, suffix)
    if (dir.exists(d)) {
      unlink(d, recursive = TRUE)
      message("Cleaned up ", d)
    }
  }
}
```

## Objective

Runs, in order, every R notebook of the pipeline: the questionnaire
preprocessing and merge (`code/impressions/`), then the statistical
analysis (`code/Statistics/`).

**Run this after `run_movement.py`**: `02_merge_data.Rmd` needs
`data/movement/05_qom.csv`, which only the Python pipeline produces.

Open this file in RStudio and click **Run All** (Ctrl+Alt+R), or **Knit**.
Each step renders its own notebook (with its own HTML output next to it);
if a step fails, the error from that notebook is shown here and nothing
further runs.

```{r run-questionnaire}
message("=== 1/3: 01_questionnaire_long.Rmd ===")
render_and_clean("code/impressions/01_questionnaire_long.Rmd", quiet = FALSE)
```

```{r run-merge}
message("=== 2/3: 02_merge_data.Rmd ===")
render_and_clean("code/impressions/02_merge_data.Rmd", quiet = FALSE)
```

```{r run-stats}
message("=== 3/3: statistical_analysis.Rmd ===")
render_and_clean("code/Statistics/statistical_analysis.Rmd", quiet = FALSE)
```

```{r done}
message(
  "Done. Processed data in data/imp/ and data/merged/; results organised ",
  "by hypothesis under results/{descriptive,H1,H2,H3,H4,supplementary}/."
)
```
