---
title: "First impressions and movement in schizophrenia"
subtitle: "Statistical analysis pipeline"
author: "Statistical analysis report"
date: "`r format(Sys.Date(), '%Y-%m-%d')`"
output:
  html_document:
    toc: true
    toc_float: true
    toc_depth: 3
    number_sections: true
    theme: flatly
    df_print: paged
    code_folding: show
editor_options: 
  markdown: 
    wrap: 72
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(
  echo       = TRUE,
  message    = TRUE,
  warning    = TRUE,
  fig.align  = "center",
  fig.retina = 2,
  dpi        = 300
)
```

# Introduction

This notebook reports the statistical analysis of first impressions
formed about individuals with schizophrenia (SZ) and matched healthy
controls (P), and the relationship between these impressions and the
**quantity of body movement** (QoM) of the hands and of the head,
extracted from short video stimuli.

The movement analysis uses four directional quantity-of-motion (QoM)
features, extracted at the stimulus level:

-   `move_hands_horizontal_qom_rate` — hand movement, horizontal (per second);
-   `move_hands_vertical_qom_rate` — hand movement, vertical (per second);
-   `move_head_horizontal_qom_rate` — head movement, horizontal (per second);
-   `move_head_vertical_qom_rate` — head movement, vertical (per second).

These are first correlated, then reduced to a single overall movement
component by PCA, which is used in the group and impression analyses.

------------------------------------------------------------------------

# Materials and methods

## Software

```{r software-info}
# Printed in full at the end of the report (session info), as recommended by
# SAMPL: "Name the statistical software package used in the analysis."
R.version.string
```

Analyses were performed in **R** with the following packages: `lme4`,
`lmerTest`, `emmeans`, `performance`, `effectsize`, `psych`, `car`,
`flextable`, `gt`, `ggplot2`, `patchwork`. The full session information
is reported at the end of this document.

## Statistical analyses

**Alpha level.** All inferential tests used a two-sided α = 0.05 unless
otherwise specified.

**Descriptive statistics.** Continuous variables are summarised as mean
(SD) when approximately normally distributed and as median [Q1, Q3]
otherwise. Sample sizes (N) are reported for every analysis. Standard
deviations (not standard errors) are used to convey variability.

**Group comparisons (P vs. SZ) at the stimulus level.** For each
continuous variable, the test was selected automatically: (i)
Shapiro–Wilk per group tested normality and Levene's test tested
homogeneity of variances; (ii) when both assumptions held, an
independent-samples Student *t*-test was used; (iii) when variances
differed, Welch's *t*-test; (iv) when normality failed, the Wilcoxon
rank-sum test. Effect sizes are Cohen's *d* for *t*-tests and the
rank-biserial correlation for Wilcoxon tests, each with a 95 % CI.

**Linear mixed models (LMM).** For analyses on the evaluation-level data
(one row per observer × video), LMMs were fitted with `lmerTest` using
restricted maximum likelihood (REML). Random intercepts were specified
for observer and for video stimulus,
`outcome ~ predictor(s) + (1 | obs_id) + (1 | vid)`. Denominator degrees
of freedom and *p*-values were obtained via Satterthwaite's
approximation. For every LMM we report (a) the fixed-effect estimates
with 95 % CI, (b) the marginal *R²* (variance explained by the fixed
effects) and conditional *R²* (fixed + random effects) following
Nakagawa & Schielzeth, and (c) an explicit **partition of the random
variance**: the variance component and the proportion of the residual
(unexplained) variance attributable to each random intercept (observer,
video) and to the residual, together with the intraclass correlation
(ICC = proportion of residual variance captured by the random
intercepts). The additional variance accounted for by the random
structure beyond the fixed effects is summarised by conditional *R²* −
marginal *R²*. For the two-group models (total score and each item) a
standardised mean-difference effect size (Cohen's *d*, SZ − P) is
additionally obtained from the estimated marginal means with
`emmeans::eff_size()`. Model assumptions (normality of residuals,
homoscedasticity, normality of the random-effect BLUPs, singularity and
convergence) were checked numerically and visually with
`performance::check_model()`.

**Aggregating the movement features.** The four directional QoM features
(hands and head × horizontal and vertical) were first inter-correlated at
the stimulus level (Pearson's *r* when all features were approximately
normal by Shapiro-Wilk, otherwise Spearman's *ρ*). Given their strong
inter-correlations, they were reduced by a principal component analysis on
the z-standardised features. Sampling adequacy was checked (Kaiser-Meyer-
Olkin measure; Bartlett's test of sphericity); the number of components
was determined by the Kaiser criterion (eigenvalue > 1) and parallel
analysis, and a varimax (orthogonal) rotation was applied when more than
one component was retained (no effect for a single component).
Eigenvalues, communalities and a scree plot are reported. The dominant
component — oriented so that higher scores indicate more movement and
rescaled to unit SD — was used as a single *overall body movement* score
in the subsequent analyses.

**Group difference in movement.** The group difference in overall body
movement was tested on this aggregate score at the stimulus level, with
the test auto-selected from normality and variance-homogeneity checks and
reported with its effect size.

**Group and movement jointly.** The first-impression total-score model
was then re-fitted with both diagnostic group and overall movement as
fixed effects. This estimates the movement effect adjusting for group and
quantifies the attenuation of the group difference once movement is
accounted for (a mediation-style comparison, reported descriptively).
Because group and movement are correlated at the stimulus level, their
collinearity (VIF) is reported alongside.

**Robustness.** Because several impression models showed mild
heteroscedasticity, the principal fixed effects were re-estimated with a
non-parametric **cluster bootstrap** (resampling whole video stimuli with
replacement), and compared with the model-based estimates as a
sensitivity analysis. The bootstrap is valid for the crossed observer ×
video random structure, for which sandwich-type cluster-robust estimators
are not defined.

**Correlations.** Spearman's *ρ* was used throughout (no normality
requirement, robust to the small SZ sample), with 95 % CIs from Fisher's
*z* transformation. Correlations among clinical variables are flagged as
exploratory (post-hoc) per SAMPL.

**Multiple comparisons.** When several tests addressed one family of
hypotheses, Benjamini–Hochberg (BH) false-discovery-rate correction was
applied and the adjusted *p*-values are reported alongside the raw
values (impression items: family of ten items; exploratory correlations:
family of features within each clinical variable).

**Missing data.** No imputation was performed; each test uses complete
cases on the variables involved, and sample sizes are reported per
analysis.

------------------------------------------------------------------------

# Constants, helper functions and data preparation

## Packages and constants

```{r packages-constants}
suppressPackageStartupMessages({
  library(dplyr); library(tidyr); library(readr); library(stringr)
  library(purrr); library(tibble); library(ggplot2)
  library(lme4);  library(lmerTest); library(emmeans); library(performance)
  library(broom); library(broom.mixed); library(car); library(effectsize)
  library(psych); library(flextable); library(gt); library(officer)
  library(scales); library(patchwork)
})

# ---- Colours (PLOS-One-style figures) --------------------------------------
GROUP_COLOURS <- c(
  "P"  = "#f6a7a7",   # healthy controls — pink
  "SZ" = "#9caffe"    # patients          — blue
)

# ---- Movement variables (only the two total-QoM features) ------------------
MOVEMENT_VARS <- c(
  "move_hands_horizontal_qom_rate",  # hand movement, horizontal (QoM rate)
  "move_hands_vertical_qom_rate",    # hand movement, vertical (QoM rate)
  "move_head_horizontal_qom_rate",   # head movement, horizontal (QoM rate)
  "move_head_vertical_qom_rate"      # head movement, vertical (QoM rate)
)

# Pretty labels reused in tables and figures
MOVEMENT_LABELS <- c(
  move_hands_horizontal_qom_rate = "Hands, horizontal (QoM)",
  move_hands_vertical_qom_rate   = "Hands, vertical (QoM)",
  move_head_horizontal_qom_rate  = "Head, horizontal (QoM)",
  move_head_vertical_qom_rate    = "Head, vertical (QoM)"
)

IMPRESSION_ITEMS <- c(
  "imp_sympathique", "imp_bizarre_inversed", "imp_intelligente",
  "imp_appreciable", "imp_confiance", "imp_dominante",
  "imp_conversation", "imp_temps", "imp_voisin", "imp_assis"
)

ALPHA_LEVEL <- 0.05

# ---- Output paths -----------------------------------------------------------
PATH_DATA_RAW   <- "../data/raw"
PATH_DATA_PROC  <- "../data/processed"
PATH_TABLES     <- "../results/tables"
PATH_FIGURES    <- "../results/figures"
for (p in c(PATH_DATA_PROC, PATH_TABLES, PATH_FIGURES)) {
  dir.create(p, recursive = TRUE, showWarnings = FALSE)
}
```

## Formatting helpers

```{r fmt-helpers}
# ---- p-values (APA-style, 3 decimals, "<.001" when very small) -------------
fmt_p <- function(p, digits = 3) {
  vapply(p, function(x) {
    if (is.na(x))     return("\u2014")
    if (x < 0.001)    return("<.001")
    sub("^0", "", sprintf(paste0("%.", digits, "f"), x))
  }, character(1))
}

# ---- Generic numbers --------------------------------------------------------
fmt_num <- function(x, digits = 2) {
  ifelse(is.na(x), "\u2014", sprintf(paste0("%.", digits, "f"), x))
}

# ---- 95 % CI ----------------------------------------------------------------
fmt_ci <- function(lo, hi, digits = 2) {
  ifelse(is.na(lo) | is.na(hi), "\u2014",
         sprintf("[%s, %s]", fmt_num(lo, digits), fmt_num(hi, digits)))
}

# ---- mean (SD) per SAMPL ----------------------------------------------------
fmt_msd <- function(m, s, digits = 2) {
  ifelse(is.na(m), "\u2014",
         sprintf("%s (%s)", fmt_num(m, digits), fmt_num(s, digits)))
}

# ---- median [Q1, Q3] for non-normal data ------------------------------------
fmt_med_iqr <- function(x, digits = 2) {
  if (all(is.na(x))) return("\u2014")
  q <- quantile(x, c(0.25, 0.5, 0.75), na.rm = TRUE)
  sprintf("%s [%s, %s]",
          fmt_num(q[2], digits), fmt_num(q[1], digits), fmt_num(q[3], digits))
}

# ---- Normalise emmeans CI column names (version-proof) ----------------------
normalize_emm_ci <- function(df) {
  nms <- names(df)
  if ("asymp.LCL" %in% nms && !"lower.CL" %in% nms) df$lower.CL <- df$asymp.LCL
  if ("asymp.UCL" %in% nms && !"upper.CL" %in% nms) df$upper.CL <- df$asymp.UCL
  df
}

# ---- Significance stars (used only inside figures, not in tables) ----------
p_stars <- function(p) {
  dplyr::case_when(
    is.na(p)   ~ "",
    p < 0.001  ~ "***",
    p < 0.01   ~ "**",
    p < 0.05   ~ "*",
    TRUE       ~ "ns"
  )
}

# ---- Strip the move_ / imp_ / stim_ prefix for display ---------------------
pretty_label <- function(x) {
  out <- ifelse(x %in% names(MOVEMENT_LABELS),
                unname(MOVEMENT_LABELS[x]),
                NA_character_)
  generic <- x %>%
    stringr::str_remove("^(move_|imp_|stim_)") %>%
    stringr::str_replace_all("_", " ") %>%
    stringr::str_to_sentence()
  ifelse(is.na(out), generic, out)
}

`%||%` <- function(x, y) if (is.null(x) || length(x) == 0) y else x
```

## Theme and figure-saving helpers

```{r theme-helpers}
# ---- PLOS-One-style minimal theme ------------------------------------------
theme_plos <- function(base_size = 11, base_family = "") {
  theme_minimal(base_size = base_size, base_family = base_family) %+replace%
    theme(
      panel.grid.major.x = element_blank(),
      panel.grid.minor   = element_blank(),
      panel.grid.major.y = element_line(colour = "grey90", linewidth = 0.3),
      axis.line.x        = element_line(colour = "grey25", linewidth = 0.4),
      axis.line.y        = element_line(colour = "grey25", linewidth = 0.4),
      axis.ticks         = element_line(colour = "grey25", linewidth = 0.3),
      axis.title         = element_text(size = base_size, colour = "grey15"),
      axis.text          = element_text(colour = "grey20", size = base_size - 1),
      plot.title         = element_text(size = base_size + 1, hjust = 0,
                                        colour = "grey10", margin = margin(b = 6)),
      plot.subtitle      = element_text(size = base_size - 1, hjust = 0,
                                        colour = "grey35", margin = margin(b = 8)),
      plot.caption       = element_text(size = base_size - 2, hjust = 0,
                                        colour = "grey30", lineheight = 1.2,
                                        margin = margin(t = 8)),
      strip.background   = element_rect(fill = "white", colour = NA),
      strip.text         = element_text(face = "bold", size = base_size,
                                        colour = "grey15"),
      legend.position    = "bottom",
      legend.title       = element_text(size = base_size - 1),
      legend.text        = element_text(size = base_size - 1),
      plot.background    = element_rect(fill = "white", colour = NA),
      panel.background   = element_rect(fill = "white", colour = NA),
      plot.margin        = margin(10, 12, 10, 10)
    )
}

# ---- Save TIFF + PNG (PLOS One: TIFF 300 dpi for submission) ---------------
save_figure <- function(plot, basename, width = 5.2, height = 4.2, dpi = 300) {
  dir.create(PATH_FIGURES, recursive = TRUE, showWarnings = FALSE)
  tiff_path <- file.path(PATH_FIGURES, paste0(basename, ".tif"))
  png_path  <- file.path(PATH_FIGURES, paste0(basename, ".png"))
  ggsave(tiff_path, plot = plot, width = width, height = height,
         dpi = dpi, compression = "lzw", bg = "white")
  ggsave(png_path,  plot = plot, width = width, height = height,
         dpi = dpi, bg = "white")
  message("Saved: ", tiff_path, " and ", png_path)
  invisible(c(tiff = tiff_path, png = png_path))
}

# ---- Add significance bracket above two bars -------------------------------
add_signif_bracket <- function(p, x1, x2, y_top, p_value,
                               tick_height_ratio = 0.02,
                               star_size = 4.5) {
  star <- p_stars(p_value)
  y_drop <- y_top * (1 - tick_height_ratio)
  y_star <- y_top * 1.012
  p +
    annotate("segment", x = x1,   xend = x2, y = y_top, yend = y_top,
             colour = "grey20", linewidth = 0.4) +
    annotate("segment", x = x1,   xend = x1, y = y_top, yend = y_drop,
             colour = "grey20", linewidth = 0.4) +
    annotate("segment", x = x2,   xend = x2, y = y_top, yend = y_drop,
             colour = "grey20", linewidth = 0.4) +
    annotate("text",    x = (x1 + x2) / 2, y = y_star, label = star,
             size = star_size, vjust = 0, colour = "grey10")
}
```

## Publication-style flextable helper

```{r flextable-helper}
# ---- PLOS-One-style flextable: bold title above, footnote(s) below ---------
plos_table <- function(df, title, footnote = NULL,
                       digits_cols = NULL) {
  ft <- flextable::flextable(df) |>
    flextable::theme_booktabs(bold_header = TRUE) |>
    flextable::set_caption(caption = title,
                           autonum  = officer::run_autonum(
                             seq_id  = "tab", pre_label = "Table ",
                             post_label = ". ", bkm = "table")) |>
    flextable::font(fontname = "Calibri", part = "all") |>
    flextable::fontsize(size = 10, part = "all") |>
    flextable::fontsize(size = 10, part = "header") |>
    flextable::bold(part = "header") |>
    flextable::align(align = "left",   part = "header") |>
    flextable::align(align = "center", part = "body") |>
    flextable::align(j = 1, align = "left", part = "all") |>
    flextable::padding(padding = 4, part = "all") |>
    flextable::border_inner_h(border = officer::fp_border(color = "grey85", width = 0.5)) |>
    flextable::hline_top(border    = officer::fp_border(color = "grey25", width = 1.2),
                         part = "header") |>
    flextable::hline_bottom(border = officer::fp_border(color = "grey25", width = 1.2),
                            part = "header") |>
    flextable::hline_bottom(border = officer::fp_border(color = "grey25", width = 1.2),
                            part = "body")

  if (!is.null(digits_cols)) {
    for (col in names(digits_cols)) {
      if (col %in% names(df))
        ft <- flextable::colformat_double(ft, j = col,
                                          digits = digits_cols[[col]])
    }
  }
  if (!is.null(footnote)) {
    ft <- flextable::add_footer_lines(ft, values = footnote)
    ft <- flextable::italic(ft, italic = TRUE, part = "footer")
    ft <- flextable::fontsize(ft, size = 9, part = "footer")
  }
  flextable::autofit(ft)
}

# ---- Export a flextable to .docx (PLOS expects cell-based Word tables) -----
save_table <- function(ft, basename) {
  dir.create(PATH_TABLES, recursive = TRUE, showWarnings = FALSE)
  doc_path  <- file.path(PATH_TABLES, paste0(basename, ".docx"))
  flextable::save_as_docx(ft, path = doc_path)
  invisible(doc_path)
}
```

## Data-preparation functions

```{r prep-functions}
# ---- Clean and rename columns of the three source files --------------------
clean_impressions <- function(d) {
  d |>
    rename(obs_id        = CASE,
           obs_age       = Age,
           obs_gender    = Gender,
           obs_pro_sante = Pro_Sante) |>
    rename_with(
      .cols = any_of(c("Sympathique", "Bizarre", "Bizarre_inversed",
                       "Intelligente", "Appreciable", "Confiance",
                       "Dominante", "Conversation", "Temps", "Voisin",
                       "Assis", "Total_score")),
      .fn = ~ paste0("imp_", tolower(.x))
    ) |>
    mutate(
      stim_id    = str_replace(vid, "^Video_", "") |> str_replace("\\.mp4$", ""),
      stim_group = case_when(Type == "Patient" ~ "SZ",
                             Type == "Healthy" ~ "P",
                             TRUE              ~ NA_character_) |>
                   factor(levels = c("P", "SZ"))
    )
}

# Keep the four directional QoM signals (hands/head x horizontal/vertical).
clean_movement <- function(d) {
  keep_signals <- c("hands_horizontal", "hands_vertical",
                    "head_horizontal",  "head_vertical")
  d |>
    rename(stim_id = participant_id,
           move_group_check = group) |>
    filter(signal %in% keep_signals) |>
    select(stim_id, move_group_check, signal, qom_rate) |>
    pivot_wider(id_cols     = c(stim_id, move_group_check),
                names_from  = signal,
                values_from = qom_rate,
                names_glue  = "qom_rate_{signal}") |>
    rename_with(.cols = starts_with("qom_rate_"),
                .fn   = ~ paste0("move_", str_remove(.x, "^qom_rate_"),
                                 "_qom_rate"))
}

clean_stimuli <- function(d) {
  d |>
    rename(stim_id                    = Participant,
           stim_code                  = ID,
           stim_group_clinical_raw    = Group,
           stim_gender                = Gender,
           stim_age                   = Age,
           stim_education             = Education,
           stim_diagnostic_duration   = diagnostic_duration,
           stim_fNART_errors          = fNART_errors,
           stim_BDI_II                = BDI_II,
           stim_MOCA                  = MOCA,
           stim_PANSS_positive        = PANSS_Positive,
           stim_PANSS_negative        = PANSS_Negative,
           stim_PANSS_psychopathology = PANSS_Psychopathology,
           stim_chlorpro_mg           = chlorpro_mg) |>
    mutate(stim_group_clinical_check = case_when(
             stim_group_clinical_raw == "CT" ~ "P",
             stim_group_clinical_raw == "SZ" ~ "SZ",
             TRUE                            ~ NA_character_) |>
             factor(levels = c("P", "SZ")))
}

check_unique_key <- function(data, key) {
  ok <- nrow(data) == length(unique(data[[key]]))
  msg <- sprintf("Key '%s': %d rows / %d unique values - %s",
                 key, nrow(data), length(unique(data[[key]])),
                 if (ok) "OK" else "DUPLICATES")
  if (ok) message(msg) else warning(msg)
  invisible(ok)
}
```

## Statistical-test functions

```{r stat-functions}
# ---- Auto-selected two-group test (SAMPL: report assumptions + effect size)
auto_group_test <- function(data, outcome, group_var = "stim_group",
                            alpha_normality = 0.05) {

  d <- data |>
    select(all_of(c(outcome, group_var))) |>
    filter(!is.na(.data[[outcome]]), !is.na(.data[[group_var]]))

  if (!is.numeric(d[[outcome]]))
    stop(sprintf("'%s' is not numeric.", outcome))

  d[[group_var]] <- droplevels(as.factor(d[[group_var]]))
  if (nlevels(d[[group_var]]) != 2)
    stop("Exactly two groups are required.")

  groups <- levels(d[[group_var]])
  desc <- d |>
    group_by(.data[[group_var]]) |>
    summarise(n      = dplyr::n(),
              mean   = mean(.data[[outcome]]),
              sd     = sd(.data[[outcome]]),
              median = median(.data[[outcome]]),
              q1     = quantile(.data[[outcome]], 0.25),
              q3     = quantile(.data[[outcome]], 0.75),
              .groups = "drop")

  # Shapiro per group
  norm <- d |>
    group_by(.data[[group_var]]) |>
    summarise(p_shapiro = if (dplyr::n() >= 3 && dplyr::n() <= 5000)
                shapiro.test(.data[[outcome]])$p.value else NA_real_,
              .groups = "drop")
  normality_ok <- all(norm$p_shapiro > alpha_normality, na.rm = TRUE)

  # Levene
  fml <- as.formula(paste(outcome, "~", group_var))
  lev <- tryCatch(car::leveneTest(fml, data = d, center = median),
                  error = function(e) NULL)
  p_levene   <- if (!is.null(lev)) lev$`Pr(>F)`[1] else NA_real_
  vars_equal <- !is.na(p_levene) && p_levene > 0.05

  # Test choice
  if (normality_ok && vars_equal) {
    test_name <- "Student t-test"
    tt <- t.test(fml, data = d, var.equal = TRUE)
    es <- effectsize::cohens_d(fml, data = d, pooled_sd = TRUE)
    stat <- unname(tt$statistic); df_val <- unname(tt$parameter); pval <- tt$p.value
    es_name <- "Cohen's d"; es_val <- es$Cohens_d
    es_lo <- es$CI_low; es_hi <- es$CI_high
    diff_est <- unname(tt$estimate[1] - tt$estimate[2])
    diff_lo  <- tt$conf.int[1]; diff_hi <- tt$conf.int[2]
  } else if (normality_ok && !vars_equal) {
    test_name <- "Welch t-test"
    tt <- t.test(fml, data = d, var.equal = FALSE)
    es <- effectsize::cohens_d(fml, data = d, pooled_sd = FALSE)
    stat <- unname(tt$statistic); df_val <- unname(tt$parameter); pval <- tt$p.value
    es_name <- "Cohen's d"; es_val <- es$Cohens_d
    es_lo <- es$CI_low; es_hi <- es$CI_high
    diff_est <- unname(tt$estimate[1] - tt$estimate[2])
    diff_lo  <- tt$conf.int[1]; diff_hi <- tt$conf.int[2]
  } else {
    test_name <- "Wilcoxon rank-sum"
    wt <- suppressWarnings(wilcox.test(fml, data = d, exact = FALSE,
                                       conf.int = TRUE))
    es <- effectsize::rank_biserial(fml, data = d)
    stat <- unname(wt$statistic); df_val <- NA_real_; pval <- wt$p.value
    es_name <- "Rank-biserial r"; es_val <- es$r_rank_biserial
    es_lo <- es$CI_low; es_hi <- es$CI_high
    diff_est <- unname(wt$estimate)
    diff_lo  <- wt$conf.int[1]; diff_hi <- wt$conf.int[2]
  }

  tibble(
    outcome      = outcome,
    g1           = groups[1],
    n_g1         = desc$n[1],  mean_g1 = desc$mean[1], sd_g1 = desc$sd[1],
    median_g1    = desc$median[1], q1_g1 = desc$q1[1], q3_g1 = desc$q3[1],
    g2           = groups[2],
    n_g2         = desc$n[2],  mean_g2 = desc$mean[2], sd_g2 = desc$sd[2],
    median_g2    = desc$median[2], q1_g2 = desc$q1[2], q3_g2 = desc$q3[2],
    shapiro_p_g1 = norm$p_shapiro[1],
    shapiro_p_g2 = norm$p_shapiro[2],
    normality_ok = normality_ok,
    levene_p     = p_levene,
    vars_equal   = vars_equal,
    test         = test_name,
    statistic    = stat,
    df           = df_val,
    p_value      = pval,
    diff_est     = diff_est,
    diff_ci_lo   = diff_lo,
    diff_ci_hi   = diff_hi,
    es_name      = es_name,
    es           = es_val,
    es_ci_lo     = es_lo,
    es_ci_hi     = es_hi
  )
}

auto_group_test_multi <- function(data, outcomes, group_var = "stim_group",
                                  bh_correction = TRUE) {
  res <- map_dfr(outcomes,
                 ~ auto_group_test(data, .x, group_var))
  if (bh_correction)
    res$p_value_BH <- p.adjust(res$p_value, method = "BH")
  res
}

# ---- Linear mixed models ---------------------------------------------------
fit_lmm <- function(data, outcome, fixed_term, random = "(1 | obs_id) + (1 | vid)") {
  fml <- as.formula(sprintf("%s ~ %s + %s", outcome, fixed_term, random))
  lmerTest::lmer(fml, data = data, REML = TRUE)
}

extract_lmm_results <- function(model, target_term) {
  fe <- summary(model)$coefficients |>
    as.data.frame() |>
    tibble::rownames_to_column("term") |>
    as_tibble() |>
    rename(estimate = Estimate, se = `Std. Error`,
           df = df, statistic = `t value`, p_value = `Pr(>|t|)`)
  ci <- tryCatch(confint(model, method = "Wald", parm = "beta_"),
                 error = function(e) NULL)
  if (!is.null(ci)) {
    ci_df <- as.data.frame(ci) |> tibble::rownames_to_column("term") |>
      as_tibble() |> rename(ci_lo = `2.5 %`, ci_hi = `97.5 %`)
    fe <- left_join(fe, ci_df, by = "term")
  }
  row <- fe |> filter(str_detect(term, target_term)) |> slice(1)
  r2  <- tryCatch(performance::r2(model), error = function(e) NULL)
  r2m <- if (!is.null(r2)) as.numeric(r2$R2_marginal)    else NA_real_
  r2c <- if (!is.null(r2)) as.numeric(r2$R2_conditional) else NA_real_
  f2  <- if (!is.na(r2m) && r2m < 1) r2m / (1 - r2m)      else NA_real_
  tibble(
    term     = row$term  %||% NA_character_,
    estimate = row$estimate %||% NA_real_,
    se       = row$se       %||% NA_real_,
    df       = row$df       %||% NA_real_,
    statistic = row$statistic %||% NA_real_,
    ci_lo    = row$ci_lo    %||% NA_real_,
    ci_hi    = row$ci_hi    %||% NA_real_,
    p_value  = row$p_value  %||% NA_real_,
    r2_marg  = r2m,
    r2_cond  = r2c,
    f2       = f2
  )
}

# ---- All fixed effects of an LMM (used for the multi-predictor model) ------
lmm_fixed_table <- function(model, drop_intercept = TRUE) {
  co <- summary(model)$coefficients |>
    as.data.frame() |>
    tibble::rownames_to_column("term") |>
    as_tibble() |>
    rename(estimate = Estimate, se = `Std. Error`,
           df = df, statistic = `t value`, p_value = `Pr(>|t|)`)
  ci <- tryCatch(confint(model, method = "Wald", parm = "beta_"),
                 error = function(e) NULL)
  if (!is.null(ci)) {
    ci_df <- as.data.frame(ci) |> tibble::rownames_to_column("term") |>
      as_tibble() |> rename(ci_lo = `2.5 %`, ci_hi = `97.5 %`)
    co <- left_join(co, ci_df, by = "term")
  } else {
    co$ci_lo <- NA_real_; co$ci_hi <- NA_real_
  }
  if (drop_intercept) co <- co |> filter(term != "(Intercept)")
  co
}

# ---- Standardised mean-difference (Cohen's d) for a two-group LMM ----------
lmm_cohens_d <- function(model, spec = "stim_group", ref = "P") {
  out <- tibble(d = NA_real_, d_lo = NA_real_, d_hi = NA_real_)
  es_df <- tryCatch({
    emm <- emmeans::emmeans(model, specs = spec)
    # Total SD = sqrt of all variance components (observer + video + residual),
    # so d is comparable to a conventional Cohen's d rather than inflated by the
    # residual SD alone.
    vc  <- as.data.frame(lme4::VarCorr(model))
    sig <- sqrt(sum(vc$vcov[is.na(vc$var2)]))
    edf <- tryCatch(stats::df.residual(model), error = function(e) Inf)
    as.data.frame(emmeans::eff_size(emm, sigma = sig, edf = edf))
  }, error = function(e) NULL)
  if (is.null(es_df) || nrow(es_df) == 0) return(out)

  lo_col <- intersect(c("lower.CL", "asymp.LCL"), names(es_df))[1]
  hi_col <- intersect(c("upper.CL", "asymp.UCL"), names(es_df))[1]
  con <- as.character(es_df$contrast[1])
  d   <- es_df$effect.size[1]
  lo  <- es_df[[lo_col]][1]
  hi  <- es_df[[hi_col]][1]

  if (grepl(paste0("^\\s*", ref, "\\b"), con)) {
    d_new  <- -d; lo_new <- -hi; hi_new <- -lo
    d <- d_new; lo <- lo_new; hi <- hi_new
  }
  tibble(d = d, d_lo = lo, d_hi = hi)
}

run_lmm_set <- function(data, outcomes, fixed_term, target_term,
                        bh_correction = TRUE, save_diagnostics = TRUE,
                        diag_prefix = "diag") {
  res <- map_dfr(outcomes, function(out) {
    m  <- fit_lmm(data, out, fixed_term)
    if (save_diagnostics) {
      dp <- file.path(PATH_FIGURES,
                      sprintf("%s_%s.png", diag_prefix, out))
      tryCatch({
        g <- performance::check_model(m)
        ggsave(dp, g, width = 10, height = 8, dpi = 200, bg = "white")
      }, error = function(e) NULL)
    }
    bind_cols(tibble(outcome = out),
              extract_lmm_results(m, target_term))
  })
  if (bh_correction)
    res$p_value_BH <- p.adjust(res$p_value, method = "BH")
  res
}

# ---- Spearman with Fisher-z 95 % CI ----------------------------------------
spearman_with_ci <- function(x, y) {
  d <- tibble(x = x, y = y) |> tidyr::drop_na()
  n <- nrow(d)
  if (n < 4) return(tibble(n = n, rho = NA_real_, ci_lo = NA_real_,
                           ci_hi = NA_real_, p_value = NA_real_))
  ct <- suppressWarnings(cor.test(d$x, d$y, method = "spearman",
                                  exact = FALSE))
  r <- unname(ct$estimate)
  z <- atanh(r); se <- 1 / sqrt(n - 3)
  lo <- tanh(z - 1.96 * se); hi <- tanh(z + 1.96 * se)
  tibble(n = n, rho = r, ci_lo = lo, ci_hi = hi, p_value = ct$p.value)
}

spearman_multi <- function(data, y, xs, bh_correction = TRUE) {
  res <- map_dfr(xs, function(x_var) {
    out <- spearman_with_ci(data[[x_var]], data[[y]])
    bind_cols(tibble(y = y, x = x_var), out)
  })
  if (bh_correction)
    res$p_value_BH <- p.adjust(res$p_value, method = "BH")
  res
}
```

## LMM diagnostics and random-effect helpers

```{r lmm-diagnostic-helpers}
# ---- Numerical assumption checks for an lmer model -------------------------
# Returns residual normality (Shapiro), homoscedasticity (auxiliary
# Breusch-Pagan-type regression of squared residuals on fitted values),
# random-effect BLUP normality, singularity and convergence flags.
check_lmm_assumptions <- function(model) {
  res <- stats::residuals(model)
  fit <- stats::fitted(model)

  shap_res <- tryCatch(
    if (length(res) >= 3 && length(res) <= 5000)
      shapiro.test(res)$p.value else NA_real_,
    error = function(e) NA_real_)

  # Breusch-Pagan-type homoscedasticity test
  bp_p <- tryCatch({
    aux <- stats::lm(I(res^2) ~ fit)
    fst <- summary(aux)$fstatistic
    if (!is.null(fst))
      stats::pf(fst[1], fst[2], fst[3], lower.tail = FALSE) else NA_real_
  }, error = function(e) NA_real_)
  bp_p <- unname(bp_p)

  re <- tryCatch(lme4::ranef(model), error = function(e) NULL)
  shap_re <- list()
  if (!is.null(re)) {
    for (g in names(re)) {
      b <- re[[g]][[1]]
      shap_re[[g]] <- tryCatch(
        if (length(b) >= 3 && length(b) <= 5000)
          shapiro.test(b)$p.value else NA_real_,
        error = function(e) NA_real_)
    }
  }

  singular <- tryCatch(lme4::isSingular(model), error = function(e) NA)
  conv_msgs <- tryCatch(model@optinfo$conv$lme4$messages,
                        error = function(e) NULL)
  non_conv  <- !is.null(conv_msgs) && length(conv_msgs) > 0

  list(
    resid_normal_p    = shap_res,
    homosced_p        = bp_p,
    homosced_violated = !is.na(bp_p) && bp_p < 0.05,
    re_normal_p       = shap_re,
    singular          = isTRUE(singular),
    non_convergence   = non_conv
  )
}

# ---- One formatted diagnostics row -----------------------------------------
fmt_diag_row <- function(diag, label_value, label_name = "Model") {
  re_obs <- diag$re_normal_p[["obs_id"]] %||% NA_real_
  re_vid <- diag$re_normal_p[["vid"]]    %||% NA_real_
  flag <- function(p) ifelse(is.na(p), "?", ifelse(p > 0.05, "ok", "!"))
  summary_code <- paste(c(
    paste0("R:", flag(diag$resid_normal_p)),
    paste0("H:", flag(diag$homosced_p)),
    paste0("RE-obs:", flag(re_obs)),
    paste0("RE-vid:", flag(re_vid)),
    if (isTRUE(diag$singular))        "singular"  else NULL,
    if (isTRUE(diag$non_convergence)) "non-conv." else NULL
  ), collapse = "; ")
  tibble::tibble(
    !!label_name := label_value,
    `Residual normality (Shapiro p)` = fmt_p(diag$resid_normal_p),
    `Homoscedasticity (BP p)`        = fmt_p(diag$homosced_p),
    `RE obs normality (p)`           = fmt_p(re_obs),
    `RE vid normality (p)`           = fmt_p(re_vid),
    `Summary`                        = summary_code
  )
}

# ---- Random-effect variance partition --------------------------------------
# Variance and SD of each random intercept and of the residual, plus the
# proportion of the total (residual + random) variance they each capture.
re_variance_table <- function(model) {
  vc <- as.data.frame(lme4::VarCorr(model))
  vc <- vc[is.na(vc$var2), , drop = FALSE]      # keep variances, drop covariances
  total <- sum(vc$vcov)
  tibble::tibble(
    grp      = vc$grp,
    Component = ifelse(vc$grp == "Residual", "Residual (within-evaluation)",
                       paste0(vc$grp, " (random intercept)")),
    Variance = vc$vcov,
    SD       = vc$sdcor,
    prop     = vc$vcov / total
  )
}

# Formatted RE-variance flextable for a single model, with ICC + R2 context.
fmt_re_table <- function(model, title, footnote = NULL) {
  rv <- re_variance_table(model)
  icc <- sum(rv$prop[rv$grp != "Residual"])
  r2  <- tryCatch(performance::r2(model), error = function(e) NULL)
  r2m <- if (!is.null(r2)) as.numeric(r2$R2_marginal)    else NA_real_
  r2c <- if (!is.null(r2)) as.numeric(r2$R2_conditional) else NA_real_

  disp <- rv |>
    transmute(
      `Random component`               = Component,
      Variance                         = fmt_num(Variance, 3),
      SD                               = fmt_num(SD, 3),
      `% of residual+random variance`  = sprintf("%.1f%%", prop * 100)
    )

  fn <- c(
    sprintf(paste0("ICC = %.3f: proportion of the residual (unexplained) ",
                   "variance captured jointly by the observer and video ",
                   "random intercepts."), icc),
    sprintf(paste0("Marginal R2 = %s (fixed effects); conditional R2 = %s ",
                   "(fixed + random). The random structure adds %s of ",
                   "explained variance beyond the fixed effects."),
            fmt_num(r2m, 3), fmt_num(r2c, 3), fmt_num(r2c - r2m, 3))
  )
  if (!is.null(footnote)) fn <- c(footnote, fn)
  plos_table(disp, title = title, footnote = fn)
}

# Compact one-row RE summary for sets of models (e.g. the 10 items).
re_proportions_row <- function(model, label, label_name = "Model") {
  rv <- re_variance_table(model)
  getp <- function(g) {
    v <- rv$prop[rv$grp == g]; if (length(v)) v[1] else NA_real_
  }
  obs_p <- getp("obs_id"); vid_p <- getp("vid"); res_p <- getp("Residual")
  tibble::tibble(
    !!label_name := label,
    `Observer var (%)` = sprintf("%.1f%%", obs_p * 100),
    `Video var (%)`    = sprintf("%.1f%%", vid_p * 100),
    `Residual var (%)` = sprintf("%.1f%%", res_p * 100),
    `ICC`              = fmt_num(obs_p + vid_p, 3)
  )
}

# ---- Sample-size / model-fit line for article-ready reporting ---------------
# Returns "N = n evaluations from k observers and m videos. AIC = ...; BIC = ..."
model_info_line <- function(model) {
  mf <- stats::model.frame(model)
  ng <- tryCatch(lme4::ngrps(model), error = function(e) c(obs_id = NA, vid = NA))
  n_obs_id <- if ("obs_id" %in% names(ng)) unname(ng["obs_id"]) else NA
  n_vid    <- if ("vid"    %in% names(ng)) unname(ng["vid"])    else NA
  sprintf(paste0("N = %d evaluations from %s observers and %s videos. ",
                 "AIC = %s; BIC = %s; REML deviance = %s."),
          nrow(mf), n_obs_id, n_vid,
          fmt_num(stats::AIC(model), 1), fmt_num(stats::BIC(model), 1),
          fmt_num(as.numeric(-2 * as.numeric(stats::logLik(model))), 1))
}

# ---- Cluster bootstrap for heteroscedasticity-robust LMM inference ----------
# Resamples whole clusters (stimuli / vid) with replacement and refits the
# model, giving robust SEs, percentile CIs and p-values. Works with CROSSED
# random effects (unlike CR2 sandwich estimators, which require nesting).
# Resampled clusters are relabelled so repeated draws form distinct RE levels.
cluster_bootstrap_lmm <- function(model, cluster = "vid", B = 1000, seed = 1) {
  set.seed(seed)
  data <- stats::model.frame(model)
  form <- stats::formula(model)
  if (!cluster %in% names(data)) return(NULL)

  idx_by   <- split(seq_len(nrow(data)), data[[cluster]])
  clusters <- names(idx_by)
  nclus    <- length(clusters)
  fe0      <- lme4::fixef(model)
  terms    <- names(fe0)
  boot     <- matrix(NA_real_, B, length(terms), dimnames = list(NULL, terms))
  ctrl     <- lme4::lmerControl(check.conv.singular = "ignore",
                                calc.derivs = FALSE)

  for (b in seq_len(B)) {
    samp <- sample(clusters, nclus, replace = TRUE)
    lens <- lengths(idx_by[samp])
    newd <- data[unlist(idx_by[samp], use.names = FALSE), , drop = FALSE]
    newd[[cluster]] <- factor(rep(seq_len(nclus), times = lens))  # distinct draws
    fit <- tryCatch(suppressWarnings(suppressMessages(
      lme4::lmer(form, data = newd, REML = TRUE, control = ctrl))),
      error = function(e) NULL)
    if (!is.null(fit)) {
      fb <- lme4::fixef(fit)
      boot[b, names(fb)] <- fb
    }
  }

  qs <- apply(boot, 2, stats::quantile, probs = c(.025, .975), na.rm = TRUE)
  tibble::tibble(
    term     = terms,
    estimate = unname(fe0[terms]),
    boot_se  = apply(boot, 2, stats::sd, na.rm = TRUE),
    ci_lo    = qs[1, ],
    ci_hi    = qs[2, ],
    p_value  = vapply(terms, function(tn) {
      v <- boot[, tn]; v <- v[!is.na(v)]
      if (!length(v)) NA_real_ else min(1, 2 * min(mean(v <= 0), mean(v >= 0)))
    }, numeric(1)),
    n_boot   = colSums(!is.na(boot))
  )
}
```

## Read and merge data

```{r read-data, eval=TRUE}
imp_raw  <- read_csv(file.path(PATH_DATA_RAW, "data_imp.csv"),
                     show_col_types = FALSE)
mov_raw  <- read_csv(file.path(PATH_DATA_RAW, "all_participants_features.csv"),
                     show_col_types = FALSE)
stim_raw <- read_delim(file.path(PATH_DATA_RAW, "stimuli_participants.csv"),
                       delim = ";", escape_double = FALSE, trim_ws = TRUE,
                       show_col_types = FALSE)

imp  <- clean_impressions(imp_raw)
mov  <- clean_movement(mov_raw)
stim <- clean_stimuli(stim_raw)

# Sanity checks
check_unique_key(mov,  "stim_id")
check_unique_key(stim, "stim_id")

# Group consistency between impression file (Type) and stimuli file (Group)
group_check <- imp |>
  distinct(stim_id, stim_group) |>
  left_join(stim |> select(stim_id, stim_group_clinical_check), by = "stim_id") |>
  filter(!is.na(stim_group_clinical_check),
         as.character(stim_group) != as.character(stim_group_clinical_check))
if (nrow(group_check) > 0) warning("Group mismatch detected; see object 'group_check'.")

# Merge into the long evaluation-level table
data_complete <- imp |>
  left_join(mov  |> select(-any_of("move_group_check")), by = "stim_id") |>
  left_join(stim |> select(-any_of(c("stim_group_clinical_raw",
                                      "stim_group_clinical_check"))),
            by = "stim_id")

# Stimulus-level table (no repeated rows)
data_stimuli <- data_complete |>
  distinct(stim_id, .keep_all = TRUE) |>
  select(stim_id, stim_group, stim_gender, stim_age, stim_education,
         stim_BDI_II, stim_fNART_errors, stim_MOCA,
         stim_PANSS_positive, stim_PANSS_negative, stim_PANSS_psychopathology,
         stim_diagnostic_duration, stim_chlorpro_mg,
         all_of(intersect(MOVEMENT_VARS, names(data_complete))))

# Observer-level table
data_observers <- data_complete |>
  distinct(obs_id, .keep_all = TRUE) |>
  select(obs_id, obs_age, obs_gender, obs_pro_sante)

# Export processed CSVs
write_csv(data_complete,  file.path(PATH_DATA_PROC, "data_complete.csv"))
write_csv(data_stimuli,   file.path(PATH_DATA_PROC, "data_stimuli.csv"))
write_csv(data_observers, file.path(PATH_DATA_PROC, "data_observers.csv"))
```
```{r}
library(dplyr)
recode_grp <- function(d) {
  d |> mutate(group_label = factor(stim_group,
                                   levels = c("P", "SZ"),
                                   labels = c("HC", "IDS")))
}
data_complete <- recode_grp(data_complete)
data_stimuli  <- recode_grp(data_stimuli)
```

------------------------------------------------------------------------

# Results

## Reliability of the first-impression scale

**Purpose.** Estimate the internal consistency of the 10
first-impression items used to compute the total impression score.

**Items used.** `imp_bizarre_inversed` is the recoded (reverse) version
of `imp_bizarre`; only the inverted version is included in the score and
in the reliability analysis.

```{r cronbach}
alpha_obj <- psych::alpha(data_complete[, IMPRESSION_ITEMS, drop = FALSE],
                          check.keys = FALSE, warnings = FALSE)

alpha_summary <- tibble(
  `N items`                       = length(IMPRESSION_ITEMS),
  `N observations`                = sum(complete.cases(
                                       data_complete[, IMPRESSION_ITEMS])),
  `Cronbach's α (raw)`            = alpha_obj$total$raw_alpha,
  `Cronbach's α (standardised)`   = alpha_obj$total$std.alpha,
  `Guttman's λ6`                  = alpha_obj$total$`G6(smc)`,
  `Mean inter-item correlation`   = alpha_obj$total$average_r
)

ft_alpha <- alpha_summary |>
  mutate(across(where(is.numeric), ~ fmt_num(.x, 3))) |>
  plos_table(
    title = "Internal consistency of the 10-item first-impression scale.",
    footnote = paste0(
      "Reliability estimated with the psych::alpha() function. ",
      "Items: ", paste(IMPRESSION_ITEMS, collapse = ", "), ". ",
      "imp_bizarre_inversed is the reverse-scored version of imp_bizarre."
    )
  )
ft_alpha
save_table(ft_alpha, "table_01_cronbach_alpha")

# Alpha if item dropped (diagnostic - reported as S1)
alpha_if_dropped <- alpha_obj$alpha.drop |>
  as.data.frame() |>
  tibble::rownames_to_column("Item") |>
  as_tibble() |>
  transmute(Item,
            `α (raw) if dropped`          = fmt_num(raw_alpha, 3),
            `α (standardised) if dropped` = fmt_num(std.alpha, 3),
            `Mean r if dropped`           = fmt_num(average_r, 3))

ft_alpha_drop <- plos_table(
  alpha_if_dropped,
  title = "Cronbach's α and mean inter-item correlation if each item is dropped.",
  footnote = "Items are not removed in subsequent analyses; this is a diagnostic."
)
ft_alpha_drop
save_table(ft_alpha_drop, "table_S1_cronbach_if_dropped")
```

## Characteristics of stimulus participants

**Purpose.** Describe and compare the two stimulus groups (P vs. SZ) on
demographic and general cognitive variables.

**Method.** Each continuous variable was tested with
`auto_group_test()`, which selects between Student's *t*-test, Welch's
*t*-test, and the Wilcoxon rank-sum test on the basis of Shapiro–Wilk
and Levene tests. Gender is reported as count and column percentage;
group differences in gender are tested with Pearson's χ² (or Fisher's
exact when expected counts are \< 5).

```{r stimuli-characteristics}
vars_stim_continu <- c("stim_age", "stim_education",
                       "stim_BDI_II", "stim_fNART_errors", "stim_MOCA")

# --- Continuous variables: auto-test ----------------------------------------
stim_tests <- auto_group_test_multi(data_stimuli, vars_stim_continu,
                                    group_var = "stim_group",
                                    bh_correction = FALSE)

# --- Gender (chi-square / Fisher) -------------------------------------------
gender_tab <- table(data_stimuli$stim_gender, data_stimuli$stim_group)
gender_test <- if (any(chisq.test(gender_tab)$expected < 5)) {
  res <- fisher.test(gender_tab)
  list(name = "Fisher's exact", stat = NA_real_,
       df = NA_real_, p = res$p.value, es = NA, lo = NA, hi = NA)
} else {
  res <- chisq.test(gender_tab)
  v   <- effectsize::cramers_v(gender_tab)
  list(name = "χ²", stat = unname(res$statistic), df = unname(res$parameter),
       p = res$p.value, es = v$Cramers_v_adjusted,
       lo = v$CI_low, hi = v$CI_high)
}

# --- Build the publication table --------------------------------------------
n_p  <- sum(data_stimuli$stim_group == "P",  na.rm = TRUE)
n_sz <- sum(data_stimuli$stim_group == "SZ", na.rm = TRUE)

gender_rows <- data_stimuli |>
  count(stim_group, stim_gender) |>
  group_by(stim_group) |>
  mutate(pct = n / sum(n) * 100) |>
  ungroup() |>
  mutate(cell = sprintf("%d (%.1f%%)", n, pct)) |>
  select(stim_group, stim_gender, cell) |>
  pivot_wider(names_from = stim_group, values_from = cell)

gender_first_row <- tibble(
  Variable    = paste0("Gender, n (%)"),
  P_value     = gender_rows[[2]][1],
  SZ_value    = gender_rows[[3]][1],
  Test        = gender_test$name,
  Statistic   = if (!is.na(gender_test$stat))
                  sprintf("%s%s",
                          fmt_num(gender_test$stat, 2),
                          ifelse(is.na(gender_test$df), "",
                                 sprintf(" (df = %d)", gender_test$df)))
                else "\u2014",
  p           = fmt_p(gender_test$p),
  `Effect size (95% CI)` = if (!is.na(gender_test$es))
    sprintf("V = %s %s", fmt_num(gender_test$es, 2),
            fmt_ci(gender_test$lo, gender_test$hi, 2))
    else "\u2014"
)

continu_rows <- stim_tests |>
  mutate(
    Variable = paste0(pretty_label(outcome), ", mean (SD)"),
    P_value  = fmt_msd(mean_g1, sd_g1, 2),
    SZ_value = fmt_msd(mean_g2, sd_g2, 2),
    Test     = test,
    Statistic = ifelse(is.na(df),
                       fmt_num(statistic, 2),
                       sprintf("%s (df = %s)",
                               fmt_num(statistic, 2), fmt_num(df, 1))),
    p        = fmt_p(p_value),
    `Effect size (95% CI)` = sprintf("%s = %s %s",
                                     ifelse(str_detect(es_name, "Cohen"), "d", "r"),
                                     fmt_num(es, 2),
                                     fmt_ci(es_ci_lo, es_ci_hi, 2))
  ) |>
  select(Variable, P_value, SZ_value, Test, Statistic, p,
         `Effect size (95% CI)`)

stim_table_df <- bind_rows(continu_rows, gender_first_row) |>
  rename(!!sprintf("P (n = %d)", n_p) := P_value,
         !!sprintf("SZ (n = %d)", n_sz) := SZ_value)

if (nrow(gender_rows) > 1) {
  for (i in 2:nrow(gender_rows)) {
    stim_table_df <- add_row(
      stim_table_df,
      Variable = paste0("  ", gender_rows$stim_gender[i]),
      !!sprintf("P (n = %d)", n_p)  := gender_rows[[2]][i],
      !!sprintf("SZ (n = %d)", n_sz) := gender_rows[[3]][i],
      Test = "", Statistic = "", p = "", `Effect size (95% CI)` = ""
    )
  }
}

ft_stim <- plos_table(
  stim_table_df,
  title = "Demographic and cognitive characteristics of stimulus participants by group.",
  footnote = c(
    paste0("P, healthy controls; SZ, individuals with a diagnosis of schizophrenia. ",
           "SD, standard deviation; CI, confidence interval; ",
           "d, Cohen's d; V, Cramér's V."),
    paste0("Tests were selected automatically based on Shapiro-Wilk (normality) ",
           "and Levene (variance homogeneity) tests; p-values are two-sided.")
  )
)
ft_stim
save_table(ft_stim, "table_02_stimuli_characteristics")
```

## SZ clinical characteristics

**Purpose.** Describe clinical characteristics of the SZ subgroup. These
variables are reported descriptively only; no group comparison applies
as they are SZ-specific.

```{r sz-clinical}
sz_only <- data_stimuli |> filter(stim_group == "SZ")
vars_sz <- c("stim_PANSS_positive", "stim_PANSS_negative",
             "stim_PANSS_psychopathology",
             "stim_diagnostic_duration", "stim_chlorpro_mg")

sz_descr <- map_dfr(vars_sz, function(v) {
  x <- sz_only[[v]]
  shap <- if (sum(!is.na(x)) >= 3) shapiro.test(x)$p.value else NA_real_
  tibble(
    Variable           = pretty_label(v),
    n                  = sum(!is.na(x)),
    `Mean (SD)`        = fmt_msd(mean(x, na.rm = TRUE), sd(x, na.rm = TRUE), 2),
    `Median [Q1, Q3]`  = fmt_med_iqr(x, 2),
    Range              = sprintf("%s - %s",
                                 fmt_num(suppressWarnings(min(x, na.rm = TRUE)), 1),
                                 fmt_num(suppressWarnings(max(x, na.rm = TRUE)), 1)),
    `Shapiro-Wilk p`   = fmt_p(shap)
  )
})

ft_sz <- plos_table(
  sz_descr,
  title = "Clinical characteristics of the SZ subgroup.",
  footnote = paste0("Q1, first quartile; Q3, third quartile. ",
                    "PANSS, Positive and Negative Syndrome Scale. ",
                    "chlorpro mg = chlorpromazine-equivalent daily dose.")
)
ft_sz
save_table(ft_sz, "table_03_SZ_clinical_descriptives")
```

## Characteristics of observers

**Purpose.** Describe the observer panel that rated the stimulus videos.

```{r observers-descriptives}
n_obs   <- nrow(data_observers)
obs_age <- data_observers$obs_age

obs_continuous_row <- tibble(
  Variable     = "Age (years)",
  n            = sum(!is.na(obs_age)),
  Summary      = fmt_msd(mean(obs_age, na.rm = TRUE),
                         sd(obs_age,   na.rm = TRUE), 1),
  Distribution = "Mean (SD)"
)


obs_cat_rows <- list(
  obs_gender    = "Gender",
  obs_pro_sante = "Health-care professional"
) |>
  imap_dfr(function(label, var) {
    data_observers |>
      filter(!is.na(.data[[var]])) |>
      count(.data[[var]]) |>
      mutate(pct = n / sum(n) * 100,
             Variable = label,
             Level    = as.character(.data[[var]]),
             Summary  = sprintf("%d (%.1f%%)", n, pct)) |>
      select(Variable, Level, Summary)
  })

ft_obs_cont <- plos_table(
  obs_continuous_row,
  title    = "Observer panel: age summary.",
  footnote = sprintf("Total N observers = %d.", n_obs)
)
ft_obs_cont
save_table(ft_obs_cont, "table_04a_observers_age")

ft_obs_cat <- plos_table(
  obs_cat_rows,
  title    = "Observer panel: gender and health-care professional status.",
  footnote = sprintf("Total N observers = %d.", n_obs)
)
ft_obs_cat
save_table(ft_obs_cat, "table_04b_observers_categorical")
```

## First impressions: descriptive statistics by group

**Purpose.** Describe the total first-impression score and each of the
ten items by group (P vs. SZ) at the evaluation level (one row per
observer × video). These are descriptive only; the inferential group
comparison is the linear mixed model in the next section.

```{r impressions-descriptives}
imp_targets <- c("imp_total_score", IMPRESSION_ITEMS)

imp_descr <- map_dfr(imp_targets, function(v) {
  d <- data_complete |> filter(!is.na(.data[[v]]))
  by_group <- d |>
    group_by(stim_group) |>
    summarise(n = dplyr::n(),
              m = mean(.data[[v]]),
              s = sd(.data[[v]]),
              med = median(.data[[v]]),
              q1  = quantile(.data[[v]], 0.25),
              q3  = quantile(.data[[v]], 0.75),
              .groups = "drop")
  p_row <- by_group |> filter(stim_group == "P")
  s_row <- by_group |> filter(stim_group == "SZ")
  tibble(
    Variable         = ifelse(v == "imp_total_score",
                              "Total impression score", pretty_label(v)),
    `P n`            = p_row$n,
    `P mean (SD)`    = fmt_msd(p_row$m, p_row$s, 2),
    `P median [IQR]` = sprintf("%s [%s, %s]",
                               fmt_num(p_row$med, 2),
                               fmt_num(p_row$q1, 2), fmt_num(p_row$q3, 2)),
    `SZ n`           = s_row$n,
    `SZ mean (SD)`   = fmt_msd(s_row$m, s_row$s, 2),
    `SZ median [IQR]`= sprintf("%s [%s, %s]",
                               fmt_num(s_row$med, 2),
                               fmt_num(s_row$q1, 2), fmt_num(s_row$q3, 2))
  )
})

ft_imp_descr <- plos_table(
  imp_descr,
  title    = "Descriptive statistics of the first-impression scores by group.",
  footnote = c(
    paste0("Evaluation-level data (one row per observer x video). ",
           "n is the number of evaluations contributing to each group."),
    paste0("Values are descriptive; the group difference is tested with a ",
           "linear mixed model accounting for the observer x video structure.")
  )
)
ft_imp_descr
save_table(ft_imp_descr, "table_05_impression_descriptives")
```

## First impressions: total score (linear mixed model)

**Purpose.** Test whether the total first-impression score differs
between the SZ and P groups while accounting for the nested structure of
the data (each observer rates several videos, each video is rated by
several observers).

**Variables.** Outcome: `imp_total_score`. Fixed effect: `stim_group`
(reference = P). Random intercepts: observer (`obs_id`) and video
(`vid`).

```{r lmm-total}
m_total <- fit_lmm(data_complete, "imp_total_score", "stim_group")
res_total <- extract_lmm_results(m_total, "stim_group")

# Standardised mean-difference effect size (Cohen's d, SZ - P)
d_total <- lmm_cohens_d(m_total, "stim_group", ref = "P")

# Adjusted means (emmeans) for plotting and reporting
emm_total <- emmeans::emmeans(m_total, ~ stim_group) |>
  as.data.frame() |> normalize_emm_ci()

lmm_total_tab <- tibble(
  Outcome     = "Total impression score",
  Comparison  = "SZ vs. P (P = reference)",
  `β (95% CI)`= sprintf("%s %s",
                        fmt_num(res_total$estimate, 2),
                        fmt_ci(res_total$ci_lo, res_total$ci_hi, 2)),
  SE          = fmt_num(res_total$se, 2),
  df          = fmt_num(res_total$df, 1),
  t           = fmt_num(res_total$statistic, 2),
  p           = fmt_p(res_total$p_value),
  `Cohen's d (95% CI)` = sprintf("%s %s",
                                 fmt_num(d_total$d, 2),
                                 fmt_ci(d_total$d_lo, d_total$d_hi, 2)),
  `R² marginal`    = fmt_num(res_total$r2_marg, 3),
  `R² conditional` = fmt_num(res_total$r2_cond, 3),
  `f²`        = fmt_num(res_total$f2, 3)
)

ft_lmm_total <- plos_table(
  lmm_total_tab,
  title = "Linear mixed model for the total first-impression score.",
  footnote = c(
    paste0("Model: imp_total_score ~ stim_group + (1 | obs_id) + (1 | vid). ",
           "REML estimation; Satterthwaite denominator df."),
    paste0("β, estimated mean difference (SZ - P); CI, confidence interval; ",
           "Cohen's d, adjusted mean difference (SZ - P) standardised by the ",
           "total SD (observer + video + residual); ",
           "R² marginal, variance explained by fixed effects (Nakagawa); ",
           "R² conditional, variance explained by fixed + random effects; ",
           "f² = R²marg / (1 - R²marg)."),
    model_info_line(m_total)
  )
)
ft_lmm_total
save_table(ft_lmm_total, "table_06_lmm_total_score")

# ---- Random-effect variance partition (share explained by random effects) --
ft_re_total <- fmt_re_table(
  m_total,
  title    = "Random-effect variance partition for the total-impression-score model.",
  footnote = "Random intercepts: observer (obs_id) and video (vid)."
)
ft_re_total
save_table(ft_re_total, "table_06b_lmm_total_random_effects")
```




```{r lmm-total-figure, fig.width=5, fig.height=5}
emm_total <- emm_total |>
  mutate(stim_group = factor(stim_group, levels = c("P", "SZ")))

y_top_total <- max(emm_total$upper.CL, na.rm = TRUE) * 1.10

fig_total <- ggplot(emm_total, aes(x = stim_group, y = emmean,
                                   fill = stim_group)) +
  geom_col(width = 0.6, colour = NA, linewidth = 0.3) +
  geom_errorbar(aes(ymin = lower.CL, ymax = upper.CL),
                width = 0.14, linewidth = 0.4, colour = "grey20") +
  scale_fill_manual(values = GROUP_COLOURS, guide = "none") +
  scale_x_discrete(labels = c("P" = "HC", "SZ" = "IDS")) +
  scale_y_continuous(expand = expansion(mult = c(0, 0.10))) +
  labs(title    = "First-impression score by group",
       x        = "Group",
       y        = "Total impression score",
       caption  = paste0(
         "Adjusted means and 95% confidence intervals from a linear mixed\n",
         "model with random intercepts for observer and video.\n",
         "P, healthy controls; SZ, individuals with a diagnosis of schizophrenia.\n ",
         "Asterisks: * p < .05, ** p < .01, *** p < .001; ns, not significant.")) +
  theme_plos()

fig_total <- add_signif_bracket(fig_total, x1 = 1, x2 = 2,
                                y_top = y_top_total,
                                p_value = res_total$p_value)
fig_total
save_figure(fig_total, "Fig1_total_impression_score", width = 5, height = 5)
```

**Fig 1. Total first-impression score by group.** Adjusted means and 95
% confidence intervals from a linear mixed model with random intercepts
for observer and video.

### Diagnostics — total impression score

```{r lmm-total-diagnostics}
diag_total <- check_lmm_assumptions(m_total)

tryCatch({
  g <- performance::check_model(m_total)
  ggsave(file.path(PATH_FIGURES, "diag_lmm_total_score.png"),
         g, width = 10, height = 8, dpi = 200, bg = "white")
}, error = function(e) NULL)

diag_total_tab <- fmt_diag_row(diag_total,
                                label_value = "Total impression score",
                                label_name  = "Outcome")
ft_diag_total <- plos_table(
  diag_total_tab,
  title    = "Diagnostic checks for the total first-impression-score linear mixed model.",
  footnote = c(
    paste0("p > .05 indicates the corresponding assumption is supported. ",
           "Tests: Shapiro-Wilk (residual and BLUP normality); ",
           "Breusch-Pagan-type auxiliary regression (homoscedasticity)."),
    paste0("Summary codes - R: residual normality; H: homoscedasticity; ",
           "RE-obs / RE-vid: random-effect normality; ",
           "singular: variance component on the boundary; ",
           "non-conv.: optimiser non-convergence."),
    paste0("Visual diagnostic panel saved as ",
           "outputs/figures/diag_lmm_total_score.png.")
  )
)
ft_diag_total
save_table(ft_diag_total, "table_S2_lmm_total_diagnostics")
```

## First impressions: item-level analyses

**Purpose.** Examine the effect of group on each of the 10
first-impression items separately. BH-adjusted *p*-values are reported
across the ten items.

```{r lmm-items}
items_summary <- run_lmm_set(
  data        = data_complete,
  outcomes    = IMPRESSION_ITEMS,
  fixed_term  = "stim_group",
  target_term = "stim_group",
  bh_correction    = TRUE,
  save_diagnostics = FALSE
)

# Standardised mean-difference effect size (Cohen's d, SZ - P) per item
items_d <- map_dfr(IMPRESSION_ITEMS, function(it) {
  m <- fit_lmm(data_complete, it, "stim_group")
  bind_cols(tibble(outcome = it), lmm_cohens_d(m, "stim_group", ref = "P"))
})
items_summary <- left_join(items_summary, items_d, by = "outcome")

items_tab <- items_summary |>
  mutate(
    Item        = pretty_label(outcome),
    `β (95% CI)`= sprintf("%s %s",
                          fmt_num(estimate, 2),
                          fmt_ci(ci_lo, ci_hi, 2)),
    SE          = fmt_num(se, 2),
    df          = fmt_num(df, 1),
    t           = fmt_num(statistic, 2),
    p           = fmt_p(p_value),
    `p (BH)`    = fmt_p(p_value_BH),
    `Cohen's d (95% CI)` = sprintf("%s %s",
                                   fmt_num(d, 2),
                                   fmt_ci(d_lo, d_hi, 2)),
    `R² marg`   = fmt_num(r2_marg, 3),
    `R² cond`   = fmt_num(r2_cond, 3),
    `f²`        = fmt_num(f2, 3)
  ) |>
  select(Item, `β (95% CI)`, SE, df, t, p, `p (BH)`,
         `Cohen's d (95% CI)`, `R² marg`, `R² cond`, `f²`)

ft_items <- plos_table(
  items_tab,
  title    = "Group effect (SZ vs. P) on each first-impression item from linear mixed models.",
  footnote = c(
    paste0("Each row corresponds to a separate LMM: item ~ stim_group + ",
           "(1 | obs_id) + (1 | vid). REML; Satterthwaite df."),
    paste0("β, estimated mean difference (SZ - P); Cohen's d, standardised ",
           "mean difference (SZ - P); p (BH), Benjamini-Hochberg adjusted ",
           "p-value across the 10 items. R² marginal and conditional are ",
           "Nakagawa estimates.")
  )
)
ft_items
save_table(ft_items, "table_07_lmm_items")
```

```{r lmm-items-figure-long, fig.width=10, fig.height=5}
emm_items <- map_dfr(IMPRESSION_ITEMS, function(it) {
  m   <- fit_lmm(data_complete, it, "stim_group")
  emm <- emmeans::emmeans(m, ~ stim_group) |> as.data.frame() |> normalize_emm_ci()
  emm$item <- pretty_label(it)
  emm
}) |> mutate(stim_group = factor(stim_group, levels = c("P", "SZ")))

star_df <- items_summary |>
  transmute(item = pretty_label(outcome), star = p_stars(p_value_BH))

y_pos <- emm_items |>
  group_by(item) |>
  summarise(y_top = max(upper.CL, na.rm = TRUE) * 1.05, .groups = "drop") |>
  left_join(star_df, by = "item")

fig_items <- ggplot(emm_items, aes(x = item, y = emmean, fill = stim_group)) +
  geom_col(position = position_dodge(width = 0.7), width = 0.6,
           colour = NA, linewidth = 0.3) +
  geom_errorbar(aes(ymin = lower.CL, ymax = upper.CL,
                    group = stim_group),
                position = position_dodge(width = 0.7),
                width = 0.18, linewidth = 0.4, colour = "grey20") +
  geom_text(data = y_pos,
            aes(x = item, y = y_top, label = star),
            inherit.aes = FALSE, size = 4, vjust = 0) +
  scale_fill_manual(values = GROUP_COLOURS, name = "Group",
                    labels = c("HC (Healthy Controls)", "IDS (Individuals with)")) +
  scale_x_discrete(labels = c("P" = "HC", "SZ" = "IDS")) +
  scale_y_continuous(expand = expansion(mult = c(0, 0.12))) +
  labs(title   = "First-impression item score by group",
       x       = "First-impression item",
       y       = "First-impression score",
       caption = paste0(
         "Adjusted means and 95% confidence intervals from linear mixed models ",
         "with random intercepts for observer and video, fitted separately for ",
         "each item.\n",
         "Asterisks indicate Benjamini-Hochberg adjusted p-values across the 10 ",
         "items (* p < .05, ** p < .01, *** p < .001; ns, not significant).")) +
  theme_plos() +
  theme(axis.text.x = element_text(angle = 30, hjust = 1))

fig_items
save_figure(fig_items, "Fig2_impression_items_by_group",
            width = 10.5, height = 5)
```

**Fig 2. Adjusted means for each first-impression item by group.**
Bars are adjusted means (emmeans) from item-level linear mixed models;
error bars are 95 % confidence intervals. Asterisks indicate
Benjamini–Hochberg adjusted *p*-values.

### Diagnostics and random effects — item-level models

```{r lmm-items-diagnostics}
items_diag_list <- map(IMPRESSION_ITEMS, function(it) {
  m    <- fit_lmm(data_complete, it, "stim_group")
  diag <- check_lmm_assumptions(m)
  list(item = it, model = m, diag = diag)
})

# 1. Assumption table (10 rows, one per item)
diag_items_tab <- map_dfr(items_diag_list, function(x) {
  fmt_diag_row(x$diag,
               label_value = pretty_label(x$item),
               label_name  = "Item")
})

ft_diag_items <- plos_table(
  diag_items_tab,
  title    = "Diagnostic checks for the 10 item-level linear mixed models.",
  footnote = c(
    paste0("p > .05 indicates the corresponding assumption is supported. ",
           "Tests: Shapiro-Wilk (residual and BLUP normality); ",
           "Breusch-Pagan-type auxiliary regression (homoscedasticity)."),
    paste0("Summary codes - R: residual normality; H: homoscedasticity; ",
           "RE-obs / RE-vid: random-effect normality; ",
           "singular: variance component on the boundary; ",
           "non-conv.: non-convergence.")
  )
)
ft_diag_items
save_table(ft_diag_items, "table_S3_lmm_items_diagnostics")

# 2. Random-effect variance partition for each item model
re_items_tab <- map_dfr(items_diag_list, function(x) {
  re_proportions_row(x$model, label = pretty_label(x$item),
                     label_name = "Item")
})
ft_re_items <- plos_table(
  re_items_tab,
  title    = "Random-effect variance partition for the 10 item-level models.",
  footnote = c(
    paste0("Each value is the percentage of the residual (unexplained) ",
           "variance captured by the observer random intercept, the video ",
           "random intercept, and the residual."),
    paste0("ICC = (observer + video) / total = proportion of variance ",
           "attributable to the random intercepts.")
  )
)
ft_re_items
save_table(ft_re_items, "table_07b_lmm_items_random_effects")
```

## Movement features: correlations

**Purpose.** Before aggregating, examine how the four directional movement
features — quantity of motion (QoM) of the hands and of the head, each in the
horizontal and vertical directions — covary across stimuli. Strong
inter-correlations justify reducing them to a single composite.

**Method.** Correlations among the four features at the stimulus level (one
value per stimulus). Pearson's *r* is used when all features are approximately
normal (Shapiro-Wilk p > .05), otherwise Spearman's *ρ*. Reporting Pearson when
appropriate keeps this step consistent with the PCA, which decomposes the
Pearson correlation matrix.

```{r movement-correlations}
mov_mat <- data_stimuli |>
  select(all_of(MOVEMENT_VARS)) |>
  filter(if_all(everything(), ~ !is.na(.)))

# Pearson if all features ~normal, otherwise Spearman
shap_p   <- vapply(mov_mat, function(z) shapiro.test(z)$p.value, numeric(1))
cor_meth <- if (all(shap_p > 0.05)) "pearson" else "spearman"
meth_lab <- if (cor_meth == "pearson") "Pearson" else "Spearman"
coef_sym <- if (cor_meth == "pearson") "r" else "rho"
cmat     <- cor(mov_mat, method = cor_meth, use = "complete.obs")

cor_tab <- as.data.frame(round(cmat, 2)) |>
  tibble::rownames_to_column("Feature") |>
  mutate(Feature = unname(MOVEMENT_LABELS[Feature]))
names(cor_tab)[-1] <- unname(MOVEMENT_LABELS[colnames(cmat)])

ft_mov_cor <- plos_table(
  cor_tab,
  title = sprintf("%s correlations among the four directional movement features.",
                  meth_lab),
  footnote = sprintf(paste0("Stimulus-level analysis (n = %d stimuli with ",
                            "complete movement data). %s coefficient (chosen ",
                            "from Shapiro-Wilk normality of all features). ",
                            "QoM, quantity of motion (rate, per second)."),
                     nrow(mov_mat), meth_lab)
)
ft_mov_cor
save_table(ft_mov_cor, "table_08_movement_correlations")
```

```{r movement-correlations-figure, fig.width=6, fig.height=5}
lev <- unname(MOVEMENT_LABELS[MOVEMENT_VARS])
cor_long <- as.data.frame(cmat) |>
  tibble::rownames_to_column("v1") |>
  pivot_longer(-v1, names_to = "v2", values_to = "coef") |>
  mutate(v1 = factor(unname(MOVEMENT_LABELS[v1]), levels = lev),
         v2 = factor(unname(MOVEMENT_LABELS[v2]), levels = rev(lev)))

fig_mov_cor <- ggplot(cor_long, aes(v1, v2, fill = coef)) +
  geom_tile(colour = "white", linewidth = 0.6) +
  geom_text(aes(label = fmt_num(coef, 2)), size = 3.4, colour = "grey15") +
  scale_fill_gradient2(low = "#3b6fb6", mid = "white", high = "#c0392b",
                       midpoint = 0, limits = c(-1, 1), name = coef_sym) +
  labs(x = NULL, y = NULL, title = "Correlations among movement features",
       caption = sprintf("Stimulus-level %s correlations.", meth_lab)) +
  theme_plos() +
  theme(axis.text.x = element_text(angle = 25, hjust = 1))
fig_mov_cor
save_figure(fig_mov_cor, "Fig3_movement_correlations", width = 6, height = 5)
```

**Fig 3. Correlations among the four directional movement features.**
Stimulus-level correlations (Pearson if features are approximately normal,
otherwise Spearman).

------------------------------------------------------------------------

## Aggregating the movement features (principal component analysis)

**Rationale.** Given the strong inter-correlations above, the four features are
reduced by PCA. Components with eigenvalue > 1 (Kaiser criterion) are retained;
if more than one is retained a **varimax** rotation is applied for
interpretability. Varimax is an *orthogonal* rotation — it keeps components
uncorrelated (PCA components already are) and clarifies the loading pattern; it
has **no effect when a single component is retained**. The dominant component
summarises overall body movement.

**Method.** PCA on the four z-standardised features at the stimulus level
(`psych::principal`), reported with sampling-adequacy diagnostics (KMO,
Bartlett), eigenvalues, communalities and a scree plot. The dominant component
is oriented so that higher = more movement, rescaled to unit SD, and merged
onto the evaluation-level data as `move_total_pc1`.

```{r movement-pca}
pca_input <- data_stimuli |>
  select(stim_id, all_of(MOVEMENT_VARS)) |>
  filter(if_all(all_of(MOVEMENT_VARS), ~ !is.na(.)))

X <- pca_input |> select(all_of(MOVEMENT_VARS))
R <- cor(X)   # Pearson matrix (what the PCA decomposes)

# --- Sampling adequacy ---
kmo  <- psych::KMO(R)
bart <- psych::cortest.bartlett(R, n = nrow(X))

# --- Eigenvalues and number of components ---
eig        <- eigen(R, only.values = TRUE)$values
n_keep     <- max(1L, sum(eig > 1))                       # Kaiser
n_parallel <- tryCatch(
  psych::fa.parallel(X, fa = "pc", plot = FALSE, n.iter = 100)$ncomp,
  error = function(e) NA_integer_)                        # parallel analysis
rot <- if (n_keep >= 2) "varimax" else "none"

pca_move <- psych::principal(X, nfactors = n_keep, rotate = rot, scores = TRUE)

# Dominant component, oriented, standardised to unit SD
ssl       <- colSums(unclass(pca_move$loadings)^2)
main_comp <- names(ssl)[which.max(ssl)]
ld_main   <- unclass(pca_move$loadings)[, main_comp]
flip      <- if (sum(ld_main) < 0) -1 else 1
score_m   <- as.numeric(pca_move$scores[, main_comp]) * flip

pca_scores <- tibble::tibble(stim_id = pca_input$stim_id,
                             move_total_pc1 = as.numeric(scale(score_m)))
data_complete <- data_complete |> left_join(pca_scores, by = "stim_id")
data_stimuli  <- data_stimuli  |> left_join(pca_scores, by = "stim_id")

# Table 1: sampling adequacy (KMO + Bartlett)
suit_tab <- tibble::tibble(
  Statistic = c("KMO (overall MSA)", "Bartlett chi-square", "Bartlett df",
                "Bartlett p", "n (stimuli)"),
  Value     = c(fmt_num(kmo$MSA, 3), fmt_num(bart$chisq, 1),
                as.character(bart$df), fmt_p(bart$p.value),
                as.character(nrow(X)))
)
ft_suit <- plos_table(
  suit_tab,
  title = "Suitability of the movement features for PCA.",
  footnote = paste0("KMO > 0.60 acceptable, > 0.80 good (per-variable MSA in ",
                    "kmo$MSAi). Bartlett's test of sphericity should be ",
                    "significant. Pearson correlation matrix.")
)
ft_suit
save_table(ft_suit, "table_09a_movement_pca_suitability")

# Table 2: eigenvalues and variance explained
eig_tab <- tibble::tibble(
  Component      = paste0("PC", seq_along(eig)),
  Eigenvalue     = fmt_num(eig, 3),
  `% variance`   = sprintf("%.1f%%", eig / sum(eig) * 100),
  `Cumulative %` = sprintf("%.1f%%", cumsum(eig) / sum(eig) * 100)
)
ft_eig <- plos_table(
  eig_tab,
  title = "Eigenvalues and variance explained.",
  footnote = sprintf(paste0("Components retained (Kaiser, eigenvalue > 1): %d; ",
                            "parallel analysis suggests: %s. Rotation: %s."),
                     n_keep, ifelse(is.na(n_parallel), "n/a", n_parallel), rot)
)
ft_eig
save_table(ft_eig, "table_09b_movement_pca_eigenvalues")

# Table 3: loadings + communalities
load_tab <- as.data.frame(round(unclass(pca_move$loadings), 3)) |>
  tibble::rownames_to_column("Feature") |>
  mutate(Feature = unname(MOVEMENT_LABELS[Feature]),
         `Communality (h2)` = fmt_num(pca_move$communality, 3))
ve <- pca_move$Vaccounted["Proportion Var", , drop = TRUE]
ft_pca <- plos_table(
  load_tab,
  title = "PCA loadings and communalities (four movement features, standardised).",
  footnote = c(
    sprintf("Variance explained: %s.",
            paste(sprintf("%s = %.1f%%", names(ve), ve * 100), collapse = "; ")),
    sprintf(paste0("Dominant component (%s), oriented so higher = more movement ",
                   "and rescaled to unit SD, used as the overall movement score. ",
                   "Communalities (h2): variance of each feature captured by the ",
                   "retained component(s); salient loadings |.| > 0.40."), main_comp)
  )
)
ft_pca
save_table(ft_pca, "table_09_movement_pca")
```

```{r movement-pca-scree, fig.width=5, fig.height=4}
fig_scree <- tibble::tibble(Component = seq_along(eig), Eigenvalue = eig) |>
  ggplot(aes(Component, Eigenvalue)) +
  geom_hline(yintercept = 1, linetype = "dashed", colour = "grey50") +
  geom_line(colour = "grey40") +
  geom_point(size = 2.8, colour = "#3b6fb6") +
  scale_x_continuous(breaks = seq_along(eig)) +
  labs(title = "Scree plot (PCA of movement features)",
       x = "Component", y = "Eigenvalue",
       caption = "Dashed line: Kaiser criterion (eigenvalue = 1).") +
  theme_plos()
fig_scree
save_figure(fig_scree, "FigS1_pca_scree", width = 5, height = 4)
```

```{r movement-pca-note, echo=FALSE, results='asis'}
if (n_keep == 1) {
  cat("A single component was retained, so no rotation applies; it captures one dimension of overall body movement, on which the four features load together.\n")
} else {
  cat(sprintf("%d components were retained and varimax-rotated; the dominant component is used as the overall movement score.\n", n_keep))
}
```

------------------------------------------------------------------------

## Movement by group

**Purpose.** Describe the four movement features by group and formally test the
group difference on the aggregate movement score (overall PCA component) at the
stimulus level.

```{r movement-by-group}
mov_summary <- map_dfr(MOVEMENT_VARS, function(v) {
  d <- data_stimuli |> filter(!is.na(.data[[v]]))
  by_group <- d |>
    group_by(stim_group) |>
    summarise(n = dplyr::n(), m = mean(.data[[v]]), s = sd(.data[[v]]),
              med = median(.data[[v]]), q1 = quantile(.data[[v]], .25),
              q3 = quantile(.data[[v]], .75),
              shap = if (dplyr::n() >= 3) shapiro.test(.data[[v]])$p.value else NA_real_,
              .groups = "drop")
  normal_p <- all(by_group$shap > 0.05, na.rm = TRUE)
  p_row <- by_group |> filter(stim_group == "P")
  s_row <- by_group |> filter(stim_group == "SZ")
  tibble(
    Feature      = unname(MOVEMENT_LABELS[v]),
    `P n`        = p_row$n,
    `P summary`  = if (normal_p) fmt_msd(p_row$m, p_row$s, 3)
                   else sprintf("%s [%s, %s]", fmt_num(p_row$med,3), fmt_num(p_row$q1,3), fmt_num(p_row$q3,3)),
    `SZ n`       = s_row$n,
    `SZ summary` = if (normal_p) fmt_msd(s_row$m, s_row$s, 3)
                   else sprintf("%s [%s, %s]", fmt_num(s_row$med,3), fmt_num(s_row$q1,3), fmt_num(s_row$q3,3)),
    Distribution = if (normal_p) "Normal" else "Non-normal"
  )
})
ft_mov_descr <- plos_table(
  mov_summary,
  title    = "Stimulus-level descriptive statistics for the four movement features, by group.",
  footnote = paste0("Mean (SD) for approximately normal variables; median ",
                    "[Q1, Q3] otherwise (Shapiro-Wilk within group, alpha = 0.05). ",
                    "QoM normalised by video duration (rate, per second).")
)
ft_mov_descr
save_table(ft_mov_descr, "table_10_movement_descriptives")

# Group difference on the aggregate movement score (stimulus level)
gt <- auto_group_test(data_stimuli, "move_total_pc1", "stim_group")
mov_grp_tab <- tibble::tibble(
  Comparison    = "P vs SZ",
  `M (SD), P`   = fmt_msd(gt$mean_g1, gt$sd_g1, 2),
  `M (SD), SZ`  = fmt_msd(gt$mean_g2, gt$sd_g2, 2),
  Test          = gt$test,
  Statistic     = fmt_num(gt$statistic, 2),
  df            = ifelse(is.na(gt$df), "\u2014", fmt_num(gt$df, 1)),
  p             = fmt_p(gt$p_value),
  `Effect size` = sprintf("%s = %s %s", gt$es_name, fmt_num(gt$es, 2),
                          fmt_ci(gt$es_ci_lo, gt$es_ci_hi, 2))
)
ft_mov_grp <- plos_table(
  mov_grp_tab,
  title    = "Group difference on the aggregate movement score (overall PCA component).",
  footnote = c(
    paste0("Stimulus-level test (one value per stimulus). The aggregate score ",
           "is standardised to unit SD; a positive effect size denotes more ",
           "movement in P than in SZ."),
    "Test auto-selected from normality (Shapiro-Wilk) and variance homogeneity (Levene)."
  )
)
ft_mov_grp
save_table(ft_mov_grp, "table_11_movement_group_difference")
```

```{r movement-by-group-figure, fig.width=5, fig.height=5}
dd_mov <- data_stimuli |>
  filter(!is.na(move_total_pc1)) |>
  mutate(stim_group = factor(stim_group, levels = c("P", "SZ")))

# Mean +/- 95% CI per group
summ_mov <- dd_mov |>
  group_by(stim_group) |>
  summarise(m = mean(move_total_pc1),
            se = sd(move_total_pc1) / sqrt(dplyr::n()),
            lo = m - 1.96 * se, hi = m + 1.96 * se, .groups = "drop")

# Significance marker from the group test computed above (gt)
gp_p <- gt$p_value
star <- if (gp_p < .001) "***" else if (gp_p < .01) "**" else
        if (gp_p < .05) "*"   else "ns"
ybar <- max(dd_mov$move_total_pc1, na.rm = TRUE) + 0.30

fig_mov_grp <- ggplot(dd_mov, aes(stim_group, move_total_pc1, fill = stim_group)) +
  geom_violin(width = 0.9, alpha = 0.55, colour = "grey45",
              linewidth = 0.4, trim = FALSE) +
  geom_jitter(width = 0.07, height = 0, size = 1.6, alpha = 0.75,
              colour = "grey25") +
  geom_errorbar(data = summ_mov,
                aes(x = stim_group, ymin = lo, ymax = hi),
                width = 0.10, linewidth = 0.6, colour = "black",
                inherit.aes = FALSE) +
  geom_point(data = summ_mov, aes(x = stim_group, y = m),
             size = 2.8, colour = "black", inherit.aes = FALSE) +
  annotate("segment", x = 1, xend = 2, y = ybar, yend = ybar, colour = "grey30") +
  annotate("text", x = 1.5, y = ybar + 0.18, label = star, size = 4.2) +
  scale_fill_manual(values = GROUP_COLOURS, guide = "none") +
  scale_x_discrete(labels = c("P" = "HC", "SZ" = "IDS")) +
  labs(title = "Overall quantity of motion by group",
       x = "Group", y = "Standardized overall quantity of motion",
       caption = paste0("Violins: distribution; points: stimuli; black: ",
                        "mean +/- 95% CI. Bracket: group test (", star, ").\n",
                        "P, healthy controls; SZ, individuals with schizophrenia.")) +
  theme_plos()
fig_mov_grp
save_figure(fig_mov_grp, "Fig4_movement_by_group", width = 5, height = 5)
```

**Fig 4. Overall body movement by group.** Stimulus-level aggregate movement
score (overall PCA component). Violins show the distribution, points the
individual stimuli, and the black marker the mean +/- 95% CI; the bracket
reports the group test.

------------------------------------------------------------------------


## First impressions predicted by overall movement (linear mixed model)

**Purpose.** Test whether overall body movement (the aggregate PCA
component) predicts the total first-impression score, with the same
observer × video random structure.

**Variables.** Outcome: `imp_total_score`. Fixed effect:
`move_total_pc1` (unit-SD overall movement score). Random intercepts:
observer (`obs_id`) and video (`vid`).

```{r lmm-movement-pc1}
m_pc1 <- fit_lmm(data_complete, "imp_total_score", "move_total_pc1")

pc1_fixed <- lmm_fixed_table(m_pc1, drop_intercept = TRUE)

r2_pc1  <- tryCatch(performance::r2(m_pc1), error = function(e) NULL)
r2m_pc1 <- if (!is.null(r2_pc1)) as.numeric(r2_pc1$R2_marginal)    else NA_real_
r2c_pc1 <- if (!is.null(r2_pc1)) as.numeric(r2_pc1$R2_conditional) else NA_real_

# Video-level R2: proportion of BETWEEN-VIDEO variance explained by movement
# (Snijders & Bosker level-2 R2: reduction in the video variance vs a model
# without movement). More informative than the marginal R2 for a video-level
# predictor, since the marginal R2 is diluted by observer + residual variance.
vid_var <- function(m) {
  vc <- as.data.frame(lme4::VarCorr(m)); vc$vcov[vc$grp == "vid"]
}
m_null_mov <- fit_lmm(data_complete, "imp_total_score", "1")   # no movement
r2_video   <- 1 - vid_var(m_pc1) / vid_var(m_null_mov)

pc1_lmm_tab <- pc1_fixed |>
  mutate(
    Predictor    = "Overall movement (PCA)",
    `β (95% CI)` = sprintf("%s %s", fmt_num(estimate, 3),
                           fmt_ci(ci_lo, ci_hi, 3)),
    SE           = fmt_num(se, 3),
    df           = fmt_num(df, 1),
    t            = fmt_num(statistic, 2),
    p            = fmt_p(p_value)
  ) |>
  select(Predictor, `β (95% CI)`, SE, df, t, p)

ft_lmm_pc1 <- plos_table(
  pc1_lmm_tab,
  title = "Linear mixed model predicting the total first-impression score from overall movement (PCA component).",
  footnote = c(
    paste0("Model: imp_total_score ~ move_total_pc1 + (1 | obs_id) + ",
           "(1 | vid). REML; Satterthwaite df."),
    sprintf(paste0("Whole-model effect sizes: marginal R2 = %s (movement), ",
                   "conditional R2 = %s (PC1 + random effects); f2 = %s."),
            fmt_num(r2m_pc1, 3), fmt_num(r2c_pc1, 3),
            fmt_num(if (!is.na(r2m_pc1) && r2m_pc1 < 1) r2m_pc1/(1-r2m_pc1) else NA_real_, 3)),
    sprintf(paste0("Video-level R2 = %s (proportion of between-video variance ",
                   "explained by movement; Snijders & Bosker level-2 R2)."),
            fmt_num(r2_video, 3)),
    paste0("PC1 is standardised to unit SD; β is the change in total ",
           "impression score per 1 SD increase in overall body movement."),
    model_info_line(m_pc1)
  )
)
ft_lmm_pc1
save_table(ft_lmm_pc1, "table_12_lmm_movement_pc1")

# ---- Random-effect variance partition --------------------------------------
ft_re_pc1 <- fmt_re_table(
  m_pc1,
  title    = "Random-effect variance partition for the PC1 movement model.",
  footnote = "Random intercepts: observer (obs_id) and video (vid)."
)
ft_re_pc1
save_table(ft_re_pc1, "table_12b_lmm_movement_pc1_random_effects")
```

```{r lmm-movement-pc1-figure, fig.width=6.5, fig.height=5}
# Fixed-effect prediction line over the PC1 range
rng_pc1 <- range(data_complete$move_total_pc1, na.rm = TRUE)
x_seq   <- seq(rng_pc1[1], rng_pc1[2], length.out = 100)
newdat  <- tibble::tibble(move_total_pc1 = x_seq)

Xp    <- model.matrix(~ move_total_pc1, newdat)
beta  <- lme4::fixef(m_pc1)
Vp    <- as.matrix(vcov(m_pc1))
predp <- as.numeric(Xp %*% beta)
sep   <- sqrt(diag(Xp %*% Vp %*% t(Xp)))
pred_pc1 <- tibble::tibble(x = x_seq, y = predp,
                           lo = predp - 1.96 * sep, hi = predp + 1.96 * sep)

# Observer-BLUP-adjusted per-video means (purged of observer tendencies)
re_obs_pc1 <- lme4::ranef(m_pc1)$obs_id |>
  tibble::rownames_to_column("obs_id") |>
  mutate(obs_id = as.character(obs_id)) |>
  rename(u_obs = `(Intercept)`)

pts_pc1 <- data_complete |>
  mutate(obs_id     = as.character(obs_id),
         stim_group = factor(stim_group, levels = c("P", "SZ"))) |>
  filter(!is.na(move_total_pc1), !is.na(imp_total_score), !is.na(obs_id)) |>
  left_join(re_obs_pc1, by = "obs_id") |>
  mutate(y_adj = imp_total_score - u_obs) |>
  group_by(stim_id, stim_group) |>
  summarise(x = dplyr::first(move_total_pc1),
            imp_adj = mean(y_adj, na.rm = TRUE), .groups = "drop")

fig_pc1 <- ggplot() +
  geom_ribbon(data = pred_pc1, aes(x = x, ymin = lo, ymax = hi),
              alpha = 0.15, fill = "grey60") +
  geom_line(data = pred_pc1, aes(x = x, y = y),
            linewidth = 0.8, colour = "black") +
  geom_point(data = pts_pc1, aes(x = x, y = imp_adj, fill = stim_group),
             shape = 21, size = 2.6, colour = "black", alpha = 0.85) +
  scale_fill_manual(
    values = GROUP_COLOURS,
    name = "Group",
    breaks = c("P", "SZ"),
    labels = c("HC", "IDS")
  ) +
  labs(x = "Overall standardized quantity of motion",
       y = "Observer total first impressions score",
       caption = paste0(
         "Black line: fixed-effect prediction from the PC1 LMM. ",
         "Grey ribbon: 95% CI. Points: observer-adjusted per-video means.\n",
         "HC, healthy controls; IDS, individuals with a diagnosis of schizophrenia.")) +
  theme_plos()
fig_pc1
save_figure(fig_pc1, "Fig5_movement_predicts_impression",
            width = 6.5, height = 5)
```

**Fig 5. Total first-impression score predicted by overall movement
(overall movement).** Black line is the fixed-effect prediction from the LMM (grey
ribbon: 95 % CI); points are observer-adjusted per-video means.

### Diagnostics — PC1 model

```{r lmm-movement-pc1-diagnostics}
diag_pc1 <- check_lmm_assumptions(m_pc1)

tryCatch({
  g <- performance::check_model(m_pc1)
  ggsave(file.path(PATH_FIGURES, "diag_lmm_movement_pc1.png"),
         g, width = 10, height = 8, dpi = 200, bg = "white")
}, error = function(e) NULL)

diag_pc1_tab <- fmt_diag_row(diag_pc1,
                             label_value = "PC1 movement model",
                             label_name  = "Model")
ft_diag_pc1 <- plos_table(
  diag_pc1_tab,
  title    = "Diagnostic checks for the PC1 movement-based first-impression model.",
  footnote = c(
    paste0("p > .05 indicates the corresponding assumption is supported. ",
           "Tests: Shapiro-Wilk (residual and BLUP normality); ",
           "Breusch-Pagan-type auxiliary regression (homoscedasticity)."),
    paste0("Summary codes - R: residual normality; H: homoscedasticity; ",
           "RE-obs / RE-vid: random-effect normality; ",
           "singular / non-conv.: boundary or convergence flags. ",
           "Visual panel: outputs/figures/diag_lmm_movement_pc1.png.")
  )
)
ft_diag_pc1
save_table(ft_diag_pc1, "table_S4_lmm_movement_diagnostics")
```

------------------------------------------------------------------------

## First impressions predicted by group and movement jointly

**Purpose.** This is the first-impressions model of the total score
(Section *total score*) with aggregate movement **added** as a second
fixed effect. It answers two questions at once: (i) does movement predict
impressions *after adjusting for diagnostic group*, and (ii) does the
group difference *shrink* once movement is taken into account — a
mediation-style question (movement as a candidate behavioural pathway for
the group difference).

**Variables.** Outcome: `imp_total_score`. Fixed effects: `stim_group`
(SZ vs P) **and** `move_total_pc1` (unit-SD aggregate movement). Random
intercepts: observer (`obs_id`) and video (`vid`).

**Caveat.** Group and movement are both stimulus-level and correlated
(movement discriminates the groups), so they are partially collinear;
the individual coefficients are therefore estimated with less precision
(VIF reported below), and the attenuation is descriptive, not a formal
mediation test.

```{r lmm-group-pc1}
m_grp_pc1 <- fit_lmm(data_complete, "imp_total_score",
                     "stim_group + move_total_pc1")

gp_fixed <- lmm_fixed_table(m_grp_pc1, drop_intercept = TRUE)

r2_gp  <- tryCatch(performance::r2(m_grp_pc1), error = function(e) NULL)
r2m_gp <- if (!is.null(r2_gp)) as.numeric(r2_gp$R2_marginal)    else NA_real_
r2c_gp <- if (!is.null(r2_gp)) as.numeric(r2_gp$R2_conditional) else NA_real_

label_term <- function(x) dplyr::case_when(
  grepl("move_total_pc1", x) ~ "Overall movement (PCA)",
  grepl("stim_group",     x) ~ "Group (SZ vs P)",
  TRUE ~ x)

gp_tab <- gp_fixed |>
  mutate(
    Predictor    = label_term(term),
    `β (95% CI)` = sprintf("%s %s", fmt_num(estimate, 2),
                           fmt_ci(ci_lo, ci_hi, 2)),
    SE           = fmt_num(se, 2),
    df           = fmt_num(df, 1),
    t            = fmt_num(statistic, 2),
    p            = fmt_p(p_value)
  ) |>
  select(Predictor, `β (95% CI)`, SE, df, t, p)

ft_lmm_gp <- plos_table(
  gp_tab,
  title = "Linear mixed model predicting the total first-impression score from group and overall movement jointly.",
  footnote = c(
    paste0("Model: imp_total_score ~ stim_group + move_total_pc1 + ",
           "(1 | obs_id) + (1 | vid). REML; Satterthwaite df. β for group is ",
           "the SZ - P difference adjusting for movement; β for PC1 is per ",
           "1 SD of movement adjusting for group."),
    sprintf(paste0("Whole-model effect sizes: marginal R2 = %s; ",
                   "conditional R2 = %s; f2 = %s."),
            fmt_num(r2m_gp, 3), fmt_num(r2c_gp, 3),
            fmt_num(if (!is.na(r2m_gp) && r2m_gp < 1) r2m_gp/(1-r2m_gp) else NA_real_, 3)),
    model_info_line(m_grp_pc1)
  )
)
ft_lmm_gp
save_table(ft_lmm_gp, "table_13_lmm_group_movement")

# ---- Attenuation of the group effect when movement is added ----------------
g_alone <- extract_lmm_results(m_total,   "stim_group")
g_adj   <- extract_lmm_results(m_grp_pc1, "stim_group")
atten_pct <- if (!is.na(g_alone$estimate) && g_alone$estimate != 0)
  (1 - g_adj$estimate / g_alone$estimate) * 100 else NA_real_

atten_tab <- tibble::tibble(
  `Group effect (SZ - P)` = c("Unadjusted (group only)",
                              "Adjusted for movement",
                              "Attenuation of |β|"),
  `β (95% CI)` = c(
    sprintf("%s %s", fmt_num(g_alone$estimate, 2),
            fmt_ci(g_alone$ci_lo, g_alone$ci_hi, 2)),
    sprintf("%s %s", fmt_num(g_adj$estimate, 2),
            fmt_ci(g_adj$ci_lo, g_adj$ci_hi, 2)),
    sprintf("%s%%", fmt_num(atten_pct, 1))),
  p = c(fmt_p(g_alone$p_value), fmt_p(g_adj$p_value), "")
)
ft_atten <- plos_table(
  atten_tab,
  title = "Attenuation of the group difference in first impressions after adjusting for aggregate movement.",
  footnote = paste0(
    "A reduction in the magnitude of the group coefficient when movement is ",
    "added is consistent with movement partly accounting for (mediating) the ",
    "group difference. Descriptive comparison, not a formal mediation test; ",
    "group and movement are correlated at the stimulus level (see below).")
)
ft_atten
save_table(ft_atten, "table_13b_group_effect_attenuation")

# ---- Collinearity between group and movement (stimulus level) --------------
g_num  <- as.integer(factor(data_complete$stim_group)) - 1L   # P = 0, SZ = 1
r_gp   <- cor(g_num, data_complete$move_total_pc1, use = "complete.obs")
vif_gp <- 1 / (1 - r_gp^2)
collin_gp_tab <- tibble::tibble(
  Quantity = c("Point-biserial r (group vs PC1)",
               "Variance inflation factor (VIF)",
               "SE inflation factor (sqrt of VIF)"),
  Value    = c(fmt_num(r_gp, 3), fmt_num(vif_gp, 2), fmt_num(sqrt(vif_gp), 2))
)
ft_collin_gp <- plos_table(
  collin_gp_tab,
  title = "Collinearity between group and aggregate movement in the joint model.",
  footnote = paste0(
    "Group and movement are correlated because movement discriminates the ",
    "groups. The VIF quantifies how much this inflates the standard errors ",
    "of the two stimulus-level predictors and tempers the interpretation of ",
    "their individual coefficients.")
)
ft_collin_gp
save_table(ft_collin_gp, "table_13c_group_movement_collinearity")

# ---- Random-effect variance partition --------------------------------------
ft_re_gp <- fmt_re_table(
  m_grp_pc1,
  title    = "Random-effect variance partition for the group + movement model.",
  footnote = "Random intercepts: observer (obs_id) and video (vid)."
)
ft_re_gp
save_table(ft_re_gp, "table_13d_lmm_group_movement_random_effects")
```

```{r lmm-group-pc1-figure, fig.width=7, fig.height=3}
pc1_row <- grepl("move_total_pc1", gp_fixed$term)
forest_df <- tibble::tibble(
  label = factor(
    c("Group (SZ - P)\nunadjusted",
      "Group (SZ - P)\nadjusting for movement",
      "Movement\nadjusting for group"),
    levels = c("Movement\nadjusting for group",
               "Group (SZ - P)\nadjusting for movement",
               "Group (SZ - P)\nunadjusted")),
  est = c(g_alone$estimate, g_adj$estimate, gp_fixed$estimate[pc1_row]),
  lo  = c(g_alone$ci_lo,    g_adj$ci_lo,    gp_fixed$ci_lo[pc1_row]),
  hi  = c(g_alone$ci_hi,    g_adj$ci_hi,    gp_fixed$ci_hi[pc1_row]),
  kind = c("Group", "Group", "Movement")
)

fig_forest <- ggplot(forest_df, aes(x = est, y = label, colour = kind)) +
  geom_vline(xintercept = 0, linetype = "dashed", colour = "grey50") +
  geom_pointrange(aes(xmin = lo, xmax = hi), linewidth = 0.7, size = 0.6) +
  scale_colour_manual(values = c(Group = GROUP_COLOURS[["SZ"]],
                                 Movement = "grey25"), guide = "none") +
  labs(x = "Effect on total first-impression score (points, 95% CI)",
       y = NULL,
       caption = paste0(
         "Movement effect adjusting for group (top); group effect with and ",
         "without adjustment for movement (middle, bottom).\n",
         "A shift of the group estimate toward zero indicates attenuation.")) +
  theme_plos()
fig_forest
save_figure(fig_forest, "Fig6_group_movement_forest", width = 7, height = 3)
```

**Fig 6. Group and movement effects on first impressions.** Point
estimates with 95 % CIs for the group difference (unadjusted and adjusted
for movement) and for movement (adjusted for group). Movement of the
group estimate toward zero indicates attenuation.

### Diagnostics — group + movement model

```{r lmm-group-pc1-diagnostics}
diag_gp <- check_lmm_assumptions(m_grp_pc1)

tryCatch({
  g <- performance::check_model(m_grp_pc1)
  ggsave(file.path(PATH_FIGURES, "diag_lmm_group_movement.png"),
         g, width = 10, height = 8, dpi = 200, bg = "white")
}, error = function(e) NULL)

diag_gp_tab <- fmt_diag_row(diag_gp, label_value = "Group + movement model",
                            label_name = "Model")
ft_diag_gp <- plos_table(
  diag_gp_tab,
  title    = "Diagnostic checks for the group + movement first-impression model.",
  footnote = c(
    paste0("p > .05 indicates the corresponding assumption is supported. ",
           "Tests: Shapiro-Wilk (residual and BLUP normality); ",
           "Breusch-Pagan-type auxiliary regression (homoscedasticity)."),
    "Visual panel: outputs/figures/diag_lmm_group_movement.png."
  )
)
ft_diag_gp
save_table(ft_diag_gp, "table_S5_lmm_group_movement_diagnostics")
```

------------------------------------------------------------------------

## Robustness: cluster bootstrap

**Purpose.** The total-score model and most item models showed mild
heteroscedasticity. To verify that the key conclusions do not depend on
the residual-variance assumptions, the principal fixed effects are
re-estimated with a **cluster bootstrap**: whole stimuli (videos) are
resampled with replacement and each model is refitted, giving robust
standard errors, percentile 95 % confidence intervals and *p*-values.
Resampling at the **stimulus** level is appropriate because group and
movement vary at that level, and — unlike CR2/sandwich estimators, which
require the random effects to be nested within the clustering variable —
the bootstrap is valid for the **crossed** observer × video structure
used here.

```{r robustness, cache=TRUE}
# Bootstrap replicates. 500 is enough for a sensitivity check; raise to
# 1000-2000 for the final run. ~B x 3 model refits, so the FIRST knit takes
# a few minutes (results are cached afterwards).
B_BOOT <- 1000

boot_total   <- cluster_bootstrap_lmm(m_total,   cluster = "vid", B = B_BOOT)
boot_pc1     <- cluster_bootstrap_lmm(m_pc1,     cluster = "vid", B = B_BOOT)
boot_grp_pc1 <- cluster_bootstrap_lmm(m_grp_pc1, cluster = "vid", B = B_BOOT)

rob_specs <- list(
  list(boot = boot_total,   model = m_total,   term = "stim_group",
       label = "Group (SZ - P): total-score model"),
  list(boot = boot_pc1,     model = m_pc1,     term = "move_total_pc1",
       label = "Overall movement: movement model"),
  list(boot = boot_grp_pc1, model = m_grp_pc1, term = "stim_group",
       label = "Group (SZ - P): group + movement"),
  list(boot = boot_grp_pc1, model = m_grp_pc1, term = "move_total_pc1",
       label = "Overall movement: group + movement")
)

rob_rows <- map_dfr(rob_specs, function(s) {
  mb <- extract_lmm_results(s$model, s$term)
  rr <- if (!is.null(s$boot)) s$boot[grepl(s$term, s$boot$term), ] else NULL
  has <- !is.null(rr) && nrow(rr) > 0
  tibble::tibble(
    Effect              = s$label,
    `β`                 = fmt_num(mb$estimate, 2),
    `Model SE`          = fmt_num(mb$se, 2),
    `Model p`           = fmt_p(mb$p_value),
    `Bootstrap SE`      = if (has) fmt_num(rr$boot_se, 2) else "—",
    `Bootstrap 95% CI`  = if (has) fmt_ci(rr$ci_lo, rr$ci_hi, 2) else "—",
    `Bootstrap p`       = if (has) fmt_p(rr$p_value) else "—"
  )
})

# Number of successful refits (surfaced so failures are visible, not silent)
n_ok <- min(c(boot_total$n_boot, boot_pc1$n_boot, boot_grp_pc1$n_boot),
            na.rm = TRUE)
boot_note <- if (all(vapply(list(boot_total, boot_pc1, boot_grp_pc1),
                            is.null, logical(1)))) {
  paste0("Bootstrap returned no result: the clustering variable 'vid' was ",
         "not found in the model frame. Check the random-effect name.")
} else {
  sprintf("Successful refits per model: at least %d of %d replicates.",
          n_ok, B_BOOT)
}

ft_robust <- plos_table(
  rob_rows,
  title = "Cluster-bootstrap sensitivity analysis for the key fixed effects.",
  footnote = c(
    sprintf(paste0("Non-parametric cluster bootstrap resampling whole video ",
                   "stimuli with replacement (%d replicates); percentile 95%% ",
                   "CIs and two-sided p-values. β is on the raw ",
                   "total-impression-score scale."), B_BOOT),
    paste0("Agreement between the model-based and bootstrap columns indicates ",
           "the conclusions are robust to the residual-variance (homoscedasticity) ",
           "assumption and to the small number of stimuli."),
    boot_note
  )
)
ft_robust
save_table(ft_robust, "table_14_robustness_cluster_bootstrap")
```

------------------------------------------------------------------------

## Exploratory correlations (SZ subgroup)

**Purpose (exploratory).** In the SZ subgroup, examine (a) the
association between the movement features and the clinical
characteristics, and (b) the association between the clinical
characteristics and the total first-impression score. Given the small SZ
sample, these analyses are exploratory (post-hoc) and use Spearman's
*ρ*.

**Method.** Spearman's *ρ* with 95 % CI from Fisher's *z*
transformation. The movement analysis uses the **overall movement score**
(aggregate PCA component); the clinical × first-impression analysis uses,
for each SZ stimulus, the mean total impression score across observers. BH
adjustment is applied across the set of clinical variables within each
analysis.

```{r exploratory-clinical-vars}
clinical_vars <- c("stim_PANSS_negative",
                   "stim_PANSS_positive",
                   "stim_PANSS_psychopathology",
                   "stim_diagnostic_duration",
                   "stim_chlorpro_mg")
```

### Movement and clinical characteristics

```{r movement-clinical-correlations}
# Re-derive the SZ subset from data_stimuli (it now contains move_total_pc1,
# added by the PCA chunk; sz_only was created earlier, before the PCA).
sz_mov <- data_stimuli |> filter(stim_group == "SZ")

cor_mov_clin <- map_dfr(clinical_vars, function(cv) {
  d <- sz_mov |> filter(!is.na(move_total_pc1), !is.na(.data[[cv]]))
  spearman_with_ci(d$move_total_pc1, d[[cv]]) |>
    mutate(y = cv)
})
cor_mov_clin$p_value_BH <- p.adjust(cor_mov_clin$p_value, method = "BH")

cor_mov_clin_tab <- cor_mov_clin |>
  mutate(
    Clinical     = pretty_label(y),
    `ρ (95% CI)` = sprintf("%s %s", fmt_num(rho, 2), fmt_ci(ci_lo, ci_hi, 2)),
    p            = fmt_p(p_value),
    `p (BH)`     = fmt_p(p_value_BH)
  ) |>
  select(Clinical, n, `ρ (95% CI)`, p, `p (BH)`)

ft_cor_mov_clin <- plos_table(
  cor_mov_clin_tab,
  title    = "Exploratory Spearman correlations between overall body movement and clinical variables (SZ subgroup).",
  footnote = c(
    paste0("SZ subgroup only (one value per stimulus). Overall movement is ",
           "the aggregate PCA component. 95% CIs from Fisher's z transformation."),
    paste0("p (BH): Benjamini-Hochberg adjusted p-value across the clinical ",
           "variables. Exploratory.")
  )
)
ft_cor_mov_clin
save_table(ft_cor_mov_clin, "table_15_movement_clinical_correlations")
```

```{r movement-clinical-figure, fig.width=9, fig.height=6}
mc_long <- sz_mov |>
  select(move_total_pc1, all_of(clinical_vars)) |>
  pivot_longer(all_of(clinical_vars), names_to = "clinical", values_to = "cval") |>
  filter(!is.na(move_total_pc1), !is.na(cval)) |>
  mutate(clinical = pretty_label(clinical))

ann_mc <- cor_mov_clin |>
  mutate(clinical = pretty_label(y),
         label    = sprintf("rho = %s, %s",
                            fmt_num(rho, 2),
                            ifelse(fmt_p(p_value_BH) == "<.001",
                                   "p (BH) <.001",
                                   paste0("p (BH) = ", fmt_p(p_value_BH)))))

fig_cor_mc <- ggplot(mc_long, aes(x = cval, y = move_total_pc1)) +
  geom_smooth(method = "lm", se = TRUE, colour = "grey20",
              fill = "grey85", linewidth = 0.6) +
  geom_point(shape = 21, fill = GROUP_COLOURS[["SZ"]], colour = "black",
             size = 2.4, alpha = 0.85) +
  geom_text(data = ann_mc, aes(x = -Inf, y = Inf, label = label),
            hjust = -0.05, vjust = 1.4, inherit.aes = FALSE,
            size = 3, colour = "grey20") +
  facet_wrap(~ clinical, scales = "free_x") +
  labs(title   = "Overall movement and clinical variables (SZ subgroup)",
       x       = "Clinical variable",
       y       = "Overall movement (PCA, SD units)",
       caption = paste0(
         "SZ subgroup. Spearman correlations (annotations: rho and BH p). ",
         "Linear fit for visualisation only.")) +
  theme_plos() +
  theme(strip.text = element_text(size = 9), axis.text = element_text(size = 8))
fig_cor_mc
save_figure(fig_cor_mc, "Fig7_movement_clinical_correlations",
            width = 9, height = 6)
```

**Fig 7. Exploratory correlations between overall body movement and
clinical variables (SZ subgroup).** Spearman correlations; annotations
show *ρ* and BH-adjusted *p*-values.

### Clinical characteristics and total first-impression score

```{r clinical-impression-correlations}
# Per-stimulus mean total impression score (SZ subgroup)
imp_by_stim <- data_complete |>
  filter(!is.na(imp_total_score)) |>
  group_by(stim_id) |>
  summarise(imp_total_mean = mean(imp_total_score), .groups = "drop")

sz_clin_imp <- sz_only |>
  left_join(imp_by_stim, by = "stim_id")

cor_clin_imp <- spearman_multi(
  sz_clin_imp, y = "imp_total_mean", xs = clinical_vars,
  bh_correction = TRUE
)

cor_clin_imp_tab <- cor_clin_imp |>
  mutate(
    Clinical     = pretty_label(x),
    n            = n,
    `ρ (95% CI)` = sprintf("%s %s", fmt_num(rho, 2), fmt_ci(ci_lo, ci_hi, 2)),
    p            = fmt_p(p_value),
    `p (BH)`     = fmt_p(p_value_BH)
  ) |>
  select(Clinical, n, `ρ (95% CI)`, p, `p (BH)`)

ft_cor_clin_imp <- plos_table(
  cor_clin_imp_tab,
  title    = "Exploratory Spearman correlations between clinical variables and the total first-impression score (SZ subgroup).",
  footnote = c(
    paste0("SZ subgroup only. The first-impression score is the per-stimulus ",
           "mean of the total impression score across observers. ",
           "95% confidence intervals from Fisher's z transformation."),
    paste0("p (BH): Benjamini-Hochberg adjusted p-value across the clinical ",
           "variables. Exploratory.")
  )
)
ft_cor_clin_imp
save_table(ft_cor_clin_imp, "table_16_clinical_impression_correlations")
```

```{r clinical-impression-figure, fig.width=11, fig.height=4.5}
scatter_ci <- sz_clin_imp |>
  select(all_of(clinical_vars), imp_total_mean) |>
  pivot_longer(all_of(clinical_vars), names_to = "clinical", values_to = "cv") |>
  mutate(clinical = pretty_label(clinical))

ann_ci <- cor_clin_imp |>
  mutate(clinical = pretty_label(x),
         label    = sprintf("rho = %s, %s",
                            fmt_num(rho, 2),
                            ifelse(fmt_p(p_value_BH) == "<.001",
                                   "p (BH) <.001",
                                   paste0("p (BH) = ", fmt_p(p_value_BH)))))

fig_cor_ci <- ggplot(scatter_ci, aes(x = cv, y = imp_total_mean)) +
  geom_point(colour = GROUP_COLOURS["SZ"], alpha = 0.85, size = 1.8) +
  geom_smooth(method = "lm", se = TRUE, colour = "grey20",
              fill = "grey85", linewidth = 0.6) +
  geom_text(data = ann_ci, aes(x = -Inf, y = Inf, label = label),
            hjust = -0.05, vjust = 1.4, inherit.aes = FALSE,
            size = 3, colour = "grey20") +
  facet_wrap(~ clinical, scales = "free_x", nrow = 1) +
  labs(title   = "Clinical variables and total first-impression score (SZ subgroup)",
       x       = "Clinical score",
       y       = "Mean total impression score (per stimulus)",
       caption = paste0(
         "Spearman correlations within the SZ subgroup. ",
         "Annotations show rho and BH-adjusted p-value. ",
         "Linear regression line shown for visualisation only.")) +
  theme_plos() +
  theme(strip.text = element_text(size = 9),
        axis.text  = element_text(size = 8))
fig_cor_ci
save_figure(fig_cor_ci, "Fig8_clinical_impression_correlations",
            width = 12, height = 4.5)
```

**Fig 8. Exploratory correlations between clinical variables and the
total first-impression score (SZ subgroup).** Spearman correlations;
annotations show *ρ* and BH-adjusted *p*-values.

------------------------------------------------------------------------

# Discussion notes

This notebook produces the descriptive and inferential outputs for the
corresponding manuscript: a publication-quality table for each analysis
(`table_01` to `table_12`, plus `S` tables), and a figure for each main
result (`Fig1` to `Fig7`). Tables are exported as Word `.docx`
(cell-based) and the underlying processed data are written to CSV in
`data/processed/`. Figures are exported as TIFF at 300 dpi and PNG for
inline viewing.

Every linear mixed model reports both the marginal/conditional *R²* and
an explicit partition of the random variance (variance components,
proportion of residual variance, ICC), so that the share of information
captured by the observer and video random effects is visible.

The interpretation of effect sizes, clinical implications, and
limitations should be developed in the manuscript itself.

------------------------------------------------------------------------

# Supporting information

**S1 Table. Cronbach's α if each item is dropped.** Internal-consistency
diagnostic of the first-impression scale. See
`results/tables/table_S1_cronbach_if_dropped.docx`.

**S2 Table. Diagnostic checks for the total first-impression-score
LMM.** Residual normality (Shapiro–Wilk), homoscedasticity
(Breusch–Pagan-type auxiliary regression), random-effect normality
(Shapiro–Wilk on the BLUPs of `obs_id` and `vid`), singular-fit and
convergence flags. See
`results/tables/table_S2_lmm_total_diagnostics.docx`.

**S3 Table. Diagnostic checks for the 10 item-level LMMs.** As S2, per
item.

**S4 Table. Diagnostic checks for the overall-movement impression
LMM.** As S2, for the model using the overall movement score (aggregate
PCA component) as a single fixed effect. See
`results/tables/table_S4_lmm_movement_diagnostics.docx`.

**S5 Table. Diagnostic checks for the group + movement first-impression
LMM.** As S2, for the model with diagnostic group and overall movement
entered jointly. See
`results/tables/table_S5_lmm_group_movement_diagnostics.docx`.

**S1 Figure. Scree plot of the movement-feature PCA.** Eigenvalues of the
components with the Kaiser reference line (eigenvalue = 1). File:
`results/figures/FigS1_pca_scree.{tiff,png}`.

**S Figures. Diagnostic plots from `performance::check_model()`.**
Residual, QQ, posterior-predictive and random-effect diagnostic panels
for each LMM. Files: `results/figures/diag_lmm_*.png`.

**S1 File. Processed long data set.** `data/processed/data_complete.csv`
— one row per observer × video evaluation.

**S2 File. Stimulus-level data set.** `data/processed/data_stimuli.csv`
— one row per stimulus.

**S3 File. Observer-level data set.**
`data/processed/data_observers.csv` — one row per observer.

------------------------------------------------------------------------

# References

1.  Bates D, Mächler M, Bolker B, Walker S. Fitting linear mixed-effects
    models using lme4. *J Stat Softw.* 2015;67(1):1–48.
2.  Kuznetsova A, Brockhoff PB, Christensen RHB. lmerTest package: tests
    in linear mixed effects models. *J Stat Softw.* 2017;82(13):1–26.
3.  Nakagawa S, Schielzeth H. A general and simple method for obtaining
    R² from generalized linear mixed-effects models. *Methods Ecol
    Evol.* 2013;4(2):133–142.
4.  Lenth RV. emmeans: estimated marginal means, aka least-squares
    means. *R package version 1.x.x.*
5.  Lüdecke D, Ben-Shachar MS, Patil I, Waggoner P, Makowski D.
    performance: an R package for assessment, comparison and testing of
    statistical models. *J Open Source Softw.* 2021;6(60):3139.
6.  Benjamini Y, Hochberg Y. Controlling the false discovery rate: a
    practical and powerful approach to multiple testing. *J R Stat Soc
    B.* 1995;57(1):289–300.
7.  Lang T, Altman D. Basic statistical reporting for articles published
    in clinical medical journals: the SAMPL Guidelines. In: Smart P,
    Maisonneuve H, Polderman A, editors. *Science Editors' Handbook.*
    European Association of Science Editors; 2013.

------------------------------------------------------------------------

# Session information

```{r session-info, message=TRUE}
sessionInfo()
```
