---
title: "1. Questionnaire preprocessing - First Impressions"
---

## Objective

This notebook:

1. Loads the questionnaire preprocessing function (`organize_data()`,
   defined in `questionnaire_preprocessing.R`);
2. Imports the raw SoSciSurvey export;
3. Filters out failed attention checks and reshapes the dataset from
   wide (one row per observer) to long format (one row per observer x
   video);
4. Exports the resulting long-format dataset for the next step
   (`02_prepare_data.Rmd`).

```{r setup, include=FALSE}
knitr::opts_chunk$set(
  echo = TRUE,
  message = FALSE,
  warning = FALSE
)
```

```{r libraries}
library(dplyr)
library(readxl)
library(readr)
```

```{r source-and-read}
source("questionnaire_preprocessing.R")
raw_data_1st_imp <- read_excel("../../data/imp/raw_data_1st_imp.xlsx")
```

```{r organize}
data_imp <- organize_data(raw_data_1st_imp)
```

```{r quick-checks}
# Quick checks
#head(data_imp)
#dim(data_imp)
#colSums(is.na(data_imp)) # Check for missing values, 0 means no missing data
```

```{r export}
out_path <- "../../data/imp/01_long.csv"
write_csv(data_imp, out_path)
if (file.exists(out_path)) {
  print(paste("Data successfully exported to", out_path))
} else {
  print("Error: Data export failed.")
}
```
