---
title: "2. Data preparation - merging questionnaire, movement and clinical data"
---

## Objective

This notebook merges the three source tables into the analysis-ready
data sets used throughout `03_statistical_analysis.Rmd`:

1. the long-format first-impression ratings
   (`data/imp/01_long.csv`, produced
   by `01_questionnaire_long.Rmd`);
2. the movement features
   (`data/movement/05_qom.csv`, produced by the
   movement pipeline, `code/movement/05_qom`);
3. the stimulus participants' clinical/demographic data
   (`data/stimuli_participants.csv`).

It renames and cleans the three sources, checks their merge keys, joins
them into 3 outputs:
- `data_complete` (one row per observer x video) — used for the first-impression LMMs (H1, H3, H4).
- `data_stimuli` (one row per video) — used for movement analyses (PCA, H2) and stimulus/clinical descriptives.
- `data_observers` (one row per observer) — used for the observer-panel descriptive table.

```{r setup, include=FALSE}
knitr::opts_chunk$set(
  echo       = TRUE,
  message    = TRUE,
  warning    = TRUE
)
```

```{r libraries}
suppressPackageStartupMessages({
  library(dplyr); library(tidyr); library(readr); library(stringr)
})

PATH_DATA   <- "../../data"
PATH_DATA_MOV   <- "../../data/movement"
PATH_DATA_IMP  <- "../../data/imp"
PATH_DATA_MERGED <- "../../data/merged"
dir.create(PATH_DATA_MOV, recursive = TRUE, showWarnings = FALSE)
```

## Data-preparation functions

```{r prep-functions}
# ---- Clean and rename columns of the three source files --------------------
clean_impressions <- function(d) {
  d |>
    rename(obs_id        = CASE,
           obs_age       = Age,
           obs_gender    = Gender,
           obs_pro_sante = Pro_Sante) |>
    rename_with(
      .cols = any_of(c("Sympathique", "Bizarre", "Bizarre_inversed",
                       "Intelligente", "Appreciable", "Confiance",
                       "Dominante", "Conversation", "Temps", "Voisin",
                       "Assis", "Total_score")),
      .fn = ~ paste0("imp_", tolower(.x))
    ) |>
    mutate(
      stim_id    = str_replace(vid, "^Video_", "") |> str_replace("\\.mp4$", ""),
      stim_group = case_when(Type == "Patient" ~ "SZ",
                             Type == "Healthy" ~ "P",
                             TRUE              ~ NA_character_) |>
                   factor(levels = c("P", "SZ"))
    )
}

# Keep the four directional QoM signals (hands/head x horizontal/vertical).
clean_movement <- function(d) {
  keep_signals <- c("hands_horizontal", "hands_vertical",
                    "head_horizontal",  "head_vertical")
  d |>
    rename(stim_id = participant_id,
           move_group_check = group) |>
    filter(signal %in% keep_signals) |>
    select(stim_id, move_group_check, signal, qom_rate) |>
    pivot_wider(id_cols     = c(stim_id, move_group_check),
                names_from  = signal,
                values_from = qom_rate,
                names_glue  = "qom_rate_{signal}") |>
    rename_with(.cols = starts_with("qom_rate_"),
                .fn   = ~ paste0("move_", str_remove(.x, "^qom_rate_"),
                                 "_qom_rate"))
}

# Note: the raw stimuli_participants.csv no longer carries a name-derived
# `ID` column (removed at the source for anonymisation), so it is not
# renamed/kept here.
clean_stimuli <- function(d) {
  d |>
    rename(stim_id                    = Participant,
           stim_group_clinical_raw    = Group,
           stim_gender                = Gender,
           stim_age                   = Age,
           stim_education             = Education,
           stim_diagnostic_duration   = diagnostic_duration,
           stim_fNART_errors          = fNART_errors,
           stim_BDI_II                = BDI_II,
           stim_MOCA                  = MOCA,
           stim_PANSS_positive        = PANSS_Positive,
           stim_PANSS_negative        = PANSS_Negative,
           stim_PANSS_psychopathology = PANSS_Psychopathology,
           stim_chlorpro_mg           = chlorpro_mg) |>
    mutate(stim_group_clinical_check = case_when(
             stim_group_clinical_raw == "CT" ~ "P",
             stim_group_clinical_raw == "SZ" ~ "SZ",
             TRUE                            ~ NA_character_) |>
             factor(levels = c("P", "SZ")))
}

check_unique_key <- function(data, key) {
  ok <- nrow(data) == length(unique(data[[key]]))
  msg <- sprintf("Key '%s': %d rows / %d unique values - %s",
                 key, nrow(data), length(unique(data[[key]])),
                 if (ok) "OK" else "DUPLICATES")
  if (ok) message(msg) else warning(msg)
  invisible(ok)
}

# Adds a readable group_label (HC / IDS) alongside stim_group (P / SZ).
recode_grp <- function(d) {
  d |> mutate(group_label = factor(stim_group,
                                   levels = c("P", "SZ"),
                                   labels = c("HC", "IDS")))
}
```

## Read, clean and merge

```{r read-data}
imp_raw  <- read_csv(file.path(PATH_DATA_IMP, "01_long.csv"),
                     show_col_types = FALSE)
mov_raw  <- read_csv(file.path(PATH_DATA_MOV, "05_qom.csv"),
                     show_col_types = FALSE)
stim_raw <- read_delim(file.path(PATH_DATA, "stimuli_participants.csv"),
                       delim = ";", escape_double = FALSE, trim_ws = TRUE,
                       show_col_types = FALSE)

imp  <- clean_impressions(imp_raw)
mov  <- clean_movement(mov_raw)
stim <- clean_stimuli(stim_raw)

# Sanity checks
check_unique_key(mov,  "stim_id")
check_unique_key(stim, "stim_id")

# Group consistency between impression file (Type) and stimuli file (Group)
group_check <- imp |>
  distinct(stim_id, stim_group) |>
  left_join(stim |> select(stim_id, stim_group_clinical_check), by = "stim_id") |>
  filter(!is.na(stim_group_clinical_check),
         as.character(stim_group) != as.character(stim_group_clinical_check))
if (nrow(group_check) > 0) warning("Group mismatch detected; see object 'group_check'.")

# Merge into the long evaluation-level table
data_complete <- imp |>
  left_join(mov  |> select(-any_of("move_group_check")), by = "stim_id") |>
  left_join(stim |> select(-any_of(c("stim_group_clinical_raw",
                                      "stim_group_clinical_check"))),
            by = "stim_id")

# Stimulus-level table (no repeated rows)
data_stimuli <- data_complete |>
  distinct(stim_id, .keep_all = TRUE) |>
  select(stim_id, stim_group, stim_gender, stim_age, stim_education,
         stim_BDI_II, stim_fNART_errors, stim_MOCA,
         stim_PANSS_positive, stim_PANSS_negative, stim_PANSS_psychopathology,
         stim_diagnostic_duration, stim_chlorpro_mg,
         all_of(intersect(c("move_hands_horizontal_qom_rate",
                             "move_hands_vertical_qom_rate",
                             "move_head_horizontal_qom_rate",
                             "move_head_vertical_qom_rate"),
                          names(data_complete))))

# Observer-level table
data_observers <- data_complete |>
  distinct(obs_id, .keep_all = TRUE) |>
  select(obs_id, obs_age, obs_gender, obs_pro_sante)

# Readable group label (HC / IDS), kept alongside stim_group (P / SZ)
data_complete <- recode_grp(data_complete)
data_stimuli  <- recode_grp(data_stimuli)
```

## Export

```{r export}
write_csv(data_complete,  file.path(PATH_DATA_MERGED, "data_complete.csv"))
write_csv(data_stimuli,   file.path(PATH_DATA_MERGED, "data_stimuli.csv"))
write_csv(data_observers, file.path(PATH_DATA_MERGED, "data_observers.csv"))

message("Wrote data_complete.csv (", nrow(data_complete), " rows), ",
        "data_stimuli.csv (", nrow(data_stimuli), " rows), ",
        "data_observers.csv (", nrow(data_observers), " rows) to ",
        PATH_DATA_MERGED)
```
