---
title: "Movement statistics"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, message = FALSE, warning = FALSE)
```

## Objective

For each participant, we compute three **global movement metrics**. Because
two participants (P10 and SZ4) had only 4 of the 6 movements recorded, a
raw mean across movements would be biased (the missing `hand_*` movements
tend to have larger values than trunk/head movements). To avoid this, each
metric is first **z-scored within each movement**, then the z-scores are
averaged across the available movements per participant:

- `mean_QoM` — standardized mean quantity of motion
- `mean_amplitude_sum` — standardized mean cumulative amplitude
- `mean_mean_speed` — standardized mean speed

Values are thus expressed in standard-deviation units around the overall
mean (0 = at average, positive = moves more than average, negative = less).

We then report descriptive statistics by group, correlations between the
three metrics, and CT vs SZ group comparisons. Tests are chosen from a
simple Shapiro-Wilk normality check: parametric when data are approximately
normal, non-parametric otherwise.

```{r libraries}
library(readr)
library(dplyr)
library(rempsyc)
library(flextable)
```

```{r setup-paths}
dir.create("../results/tables", showWarnings = FALSE, recursive = TRUE)

source("../sources/movement_statistics.R")

movement_features <- read_csv(
  "../data/all_participants_movement_features.csv",
  show_col_types = FALSE
)
```

## 1. Descriptive statistics

```{r descriptives}
movement_descriptives <- build_movement_descriptives(movement_features)

write_csv(movement_descriptives,
          "../results/tables/movement_descriptives_by_group.csv")

rempsyc::nice_table(
  movement_descriptives,
  title = c("Table 1",
            "Descriptive statistics of global movement metrics by group"),
  note  = "Values are M (SD) in standard-deviation units (metrics were z-scored within each movement before being averaged). N_CT and N_SZ report the number of participants per group."
)
```

## 2. Correlation matrix

```{r correlations}
cor_mat <- build_movement_correlation_matrix(movement_features)
method  <- attr(cor_mat, "method")

cor_df <- data.frame(Variable = rownames(cor_mat),
                     round(cor_mat, 2),
                     check.names = FALSE)

write_csv(cor_df, "../results/tables/movement_metric_correlations.csv")

rempsyc::nice_table(
  cor_df,
  title = c("Table 2",
            sprintf("Correlations between global movement metrics (%s)",
                    tools::toTitleCase(method))),
  note  = sprintf(
    "Method: %s (chosen because %s variable was %snormal in the Shapiro-Wilk test).",
    tools::toTitleCase(method),
    if (method == "pearson") "every" else "at least one",
    if (method == "pearson") "" else "non-"
  ),
  col.format.r = 2:ncol(cor_df)
)
```

## 3. Group comparisons (CT vs SZ)

```{r group-diff}
movement_group_diff <- build_movement_group_differences(movement_features)

write_csv(movement_group_diff,
          "../results/tables/movement_group_differences.csv")

rempsyc::nice_table(
  movement_group_diff,
  title = c("Table 3",
            "Group comparisons on global movement metrics"),
  note  = c("CT and SZ columns are M (SD). Statistic is t for Welch t-tests and W for Wilcoxon rank-sum tests.",
            "* p < .05, ** p < .01, *** p < .001")
)
```

```{r}
movement_plot <- plot_movement_group_differences(movement_features)
movement_plot
ggplot2::ggsave(
  filename = "../results/figures/movement_group_differences.png",
  plot = movement_plot,
  width = 9,
  height = 5.5,
  dpi = 300
)
```

