---
title: "1st impressions - Group difference on the global total"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, message = FALSE, warning = FALSE)
```

## Objective

Test whether observers rate Patients and Healthy targets differently on total first impression scores. One mixed model, APA tables, one figure.

```{r libraries}
library(readr)
library(dplyr)
library(ggplot2)
library(lme4)
library(lmerTest)
library(emmeans)
library(car)
library(rempsyc)
library(flextable)
```

```{r load-data}
dir.create("../results/tables",  showWarnings = FALSE, recursive = TRUE)
dir.create("../results/figures", showWarnings = FALSE, recursive = TRUE)

source("../sources/questionnaire_analysis.R")

data_imp <- read_csv("../data/data_imp.csv", show_col_types = FALSE)
data_imp <- prepare_questionnaire_scores(data_imp)
```

## Observer participants

```{r observers-table}
observers <- distinct(data_imp, CASE, Age, Gender)

n_obs    <- nrow(observers)
age_m    <- mean(observers$Age,  na.rm = TRUE)
age_sd   <- sd(observers$Age,    na.rm = TRUE)
n_female <- sum(observers$Gender == "F", na.rm = TRUE)

observers_df <- data.frame(
  Variable = c("Observer participants, N",
               "Age, years, M (SD)",
               "Female, n (%)"),
  Value    = c(as.character(n_obs),
               sprintf("%.2f (%.2f)", age_m, age_sd),
               sprintf("%d (%.1f%%)", n_female, 100 * n_female / n_obs))
)

rempsyc::nice_table(
  observers_df,
  title = c("Table 1", "Observer participants descriptive statistics")
)
readr::write_csv(observers_df, "../results/tables/observer_participants_descriptive_statistics.csv")
```

## Group difference on Global total

Linear mixed model: `Global_total ~ Type + num_video + (1|CASE) + (1|video_id)`. `Type` is sum-coded so the pairwise estimate from `emmeans` directly gives Patient − Healthy.

```{r model}
options(contrasts = c("contr.sum", "contr.poly"))

model_global <- lmer(
  Global_total ~ Type + num_video + (1 | CASE) + (1 | video_id),
  data = data_imp,
  REML = FALSE
)

anova_res <- car::Anova(model_global, type = 3)
emm       <- emmeans(model_global, ~ Type)
emm_df    <- as.data.frame(summary(emm))
pairs_res <- as.data.frame(summary(pairs(emm)))
```

### Estimated means per group

```{r table-emm}
emm_table <- data.frame(
  Group     = as.character(emm_df$Type),
  M         = emm_df$emmean,
  SE        = emm_df$SE,
  CI_lower  = emm_df$lower.CL,
  CI_upper  = emm_df$upper.CL
)

rempsyc::nice_table(
  emm_table,
  title = c("Table 2",
            "Estimated marginal means of Global total by target type"),
  note  = "Estimates come from a linear mixed model with random intercepts for observers (CASE) and videos (video_id), controlling for presentation order (num_video)."
)
```

### Pairwise comparison (Patient − Healthy)

```{r table-pairs}
pairs_table <- data.frame(
  Contrast = "Patient - Healthy",
  B        = pairs_res$estimate[1],
  SE       = pairs_res$SE[1],
  df       = pairs_res$df[1],
  t        = pairs_res$t.ratio[1],
  p        = pairs_res$p.value[1]
)

nice_pairs <- rempsyc::nice_table(
  pairs_table,
  title = c("Table 3",
            "Group difference on Total First Imp"),
  note  = c("B is the estimated difference between groups in raw First Imp-total units.",
            "* p < .05, ** p < .01, *** p < .001")
)
nice_pairs

# Save APA table to Word + CSV record
flextable::save_as_docx(
  nice_pairs,
  path = "../results/tables/1st_imp_group_difference.docx"
)
write_csv(pairs_table, "../results/tables/1st_imp_group_difference.csv")
```

### Figure

```{r figure, fig.width=5, fig.height=4}
p_val   <- pairs_res$p.value[1]
p_label <- dplyr::case_when(
  p_val <= .001 ~ "***",
  p_val <= .01  ~ "**",
  p_val <= .05  ~ "*",
  TRUE          ~ "ns"
)

y_max  <- max(emm_df$upper.CL)
y_line <- y_max + 0.06 * y_max
y_text <- y_max + 0.12 * y_max

fig_global <- ggplot(emm_df, aes(x = Type, y = emmean, fill = Type)) +
  geom_col(width = 0.6, color = "black", alpha = 0.9) +
  geom_errorbar(aes(ymin = lower.CL, ymax = upper.CL),
                width = 0.15, linewidth = 0.6) +
  annotate("segment", x = 1, xend = 2, y = y_line, yend = y_line) +
  annotate("text", x = 1.5, y = y_text, label = p_label, size = 6) +
  scale_fill_manual(values = c("Healthy" = "orange", "Patient" = "steelblue")) +
  labs(x = "Target type", y = "Estimated marginal mean (First Imp total)") +
  coord_cartesian(ylim = c(0, y_max + 0.22 * y_max)) +
  theme_classic(base_size = 12) +
  theme(legend.position = "none",
        axis.title = element_text(face = "bold"))

ggsave("../results/figures/1st_imp_group_difference.png",
       fig_global, width = 5, height = 4, dpi = 150)

fig_global
```
